---
title: "Spatial_analyses_plots"
author: "Trevor Tivey"
date: "2/20/2019"
output: html_document
---

```{r, libraries, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
#setwd("/local/path/to/data")
set.seed(123)

#libraries

library(data.table)
library(ggplot2)
library(spatstat)
library(plyr)
library(dplyr)
library(tidyr)
#library(aacc)


```

## Create list and data table of NN distances for aposymbiotic tentacles

```{r, apo lists, echo = FALSE}
# create list of NN csv files and combine into one large data frame

setwd("./apo_analysis/output apo/3d nn test")

#files containing NN distances of all host EdU+ nuclei to closest total nuclei
edulistapo2 = list.files(pattern = " [1-3].csv")
edulistapo2

# Create list
eduapofiles = lapply(edulistapo2, read.csv, header=T, stringsAsFactors = F)

# Create data table
eduapofiles2 <- rbindlist(eduapofiles, fill = T)
colnames(eduapofiles2) <- paste("enapo", colnames(eduapofiles2), sep = "_")


```

## Create list and data frames of NN distances for symbiotic tentacles

```{r, apo lists, echo = FALSE}

# Create list of NN csv files and combine into three large data frames for each comparison:

# 1. EdU points to nearest symbiont
# 2. Nuclei points to nearest symbiont
# 3. EdU points to nearest nuclei

setwd("./sym_analysis/output sym/3d nn test")

# lists

#files containing NN dists between host EdU+ nuclei and symbionts
edusymlist2 = list.files(pattern = "split.tif_2.csv")
edusymlist2

#files containing object coordinates of total nuclei and symbionts
nucsymlist2 = list.files(pattern = " [1-3].tif_2.csv")
nucsymlist2

#files containing object coordinates of host EdU+ nuclei and total nuclei
edunucslist2 = list.files(pattern = " [1-3].csv")
edunucslist2

# apply lists and create data tables

# Distances from EdU to symbionts
edusymfiles = lapply(edusymlist2, read.csv, header=T, stringsAsFactors = F)
edusymfiles2 <- rbindlist(edusymfiles, fill = T)
colnames(edusymfiles2) <- paste("es", colnames(edusymfiles2), sep = "_")

# Distances from nuclei to symbionts
nucsymfiles = lapply(nucsymlist2, read.csv, header=T, stringsAsFactors = F)
nucsymfiles2 <- rbindlist(nucsymfiles, fill = T)
colnames(nucsymfiles2) <- paste("ns", colnames(nucsymfiles2), sep = "_")

# Distances from EdU to nuclei
edunucsfiles = lapply(edunucslist2, read.csv, header=T, stringsAsFactors = F)
edunucsfiles2 <- rbindlist(edunucsfiles, fill = T)
colnames(edunucsfiles2) <- paste("en", colnames(edunucsfiles2), sep = "_")

```

## Test NN distance distribution for normality

```{r, apo lists, echo = FALSE}
# test for normality
qqnorm(edusymfiles2$es_dist.1)
qqline(edusymfiles2$es_dist.1)

qqnorm(nucsymfiles2$ns_dist.1)
qqline(nucsymfiles2$ns_dist.1)

qqnorm(edunucsfiles2$en_dist.1)
qqline(edunucsfiles2$en_dist.1)

qqnorm(edusymfiles2$es_dist.12)
qqline(edusymfiles2$es_dist.12)

qqnorm(nucsymfiles2$ns_dist.12)
qqline(nucsymfiles2$ns_dist.12)

qqnorm(edunucsfiles2$en_dist.12)
qqline(edunucsfiles2$en_dist.12)

qqnorm(eduapofiles2$enapo_dist.1)
qqline(eduapofiles2$enapo_dist.1)

qqnorm(eduapofiles2$enapo_dist.12)
qqline(eduapofiles2$enapo_dist.12)

# Skew in data, analyze distribution using nonparametric tests

```

## Combine all files into a single NN file for comparison

```{r, apo lists, echo = FALSE}
# Combine files

combinedfiles <- cbind(edusymfiles2, edunucsfiles2)

len <- max(nrow(combinedfiles), nrow(nucsymfiles2))
#len#delete
#nrow(combinedfiles) #delete                   
#nrow(nucsymfiles2)#delete
nrow(nucsymfiles2) - nrow(combinedfiles) # difference is 207834 
temprow <- matrix(c(rep.int(NA,length(combinedfiles))),nrow=207834,ncol=length(combinedfiles))

# Create a data.frame and give cols the same names as data
newrow <- data.frame(temprow)
colnames(newrow) <- colnames(combinedfiles)

# rbind the empty row to data
combinedsamelength <- rbind(combinedfiles,newrow)
combinedfiles2 <- cbind(combinedsamelength, nucsymfiles2)
#combinedfiles <- cbind(combinedfiles, nucfiles2) #delete


## Combine apo and sym lists

len <- max(nrow(eduapofiles2), nrow(combinedfiles2))
#len#delete
#nrow(eduapofiles2)     #delete               
#nrow(combinedfiles2)#delete
nrow(combinedfiles2) - nrow(eduapofiles2) # difference is 207653
temprow <- matrix(c(rep.int(NA,length(eduapofiles2))),nrow=207653,ncol=length(eduapofiles2))

# Create a data.frame and give cols the same names as data
newrow <- data.frame(temprow)
colnames(newrow) <- colnames(eduapofiles2)

# rbind the empty row to data
eduapofiles2samelengthapo <- rbind(eduapofiles2,newrow)

combined_nn <- cbind(eduapofiles2samelengthapo, combinedfiles2)
#combinedfiles <- cbind(combinedfiles, nucfiles2)#delete

```

## Fig 4D and 3A: Mann-Whitney U tests: is distance of EdU -> Symbiont greater than EdU --> nuclei

```{r, apo lists, echo = FALSE}

# Comparison between EdU-->symbiont distance and EdU-->host nuclei distance distributions
# EdU to symbiont distance is greater than EdU to host nuclei distances
wilcox.test(x = combined_nn$es_dist.1, y = combined_nn$en_dist.1, data = combined_nn, alternative = "greater")
wilcox.test(x = combined_nn$es_dist.2, y = combined_nn$en_dist.2, data = combined_nn, alternative = "greater")
wilcox.test(x = combined_nn$es_dist.3, y = combined_nn$en_dist.3, data = combined_nn, alternative = "greater")
wilcox.test(x = combined_nn$es_dist.4, y = combined_nn$en_dist.4, data = combined_nn, alternative = "greater")
wilcox.test(x = combined_nn$es_dist.5, y = combined_nn$en_dist.5, data = combined_nn, alternative = "greater")
wilcox.test(x = combined_nn$es_dist.6, y = combined_nn$en_dist.6, data = combined_nn, alternative = "greater")
wilcox.test(x = combined_nn$es_dist.7, y = combined_nn$en_dist.7, data = combined_nn, alternative = "greater")
wilcox.test(x = combined_nn$es_dist.8, y = combined_nn$en_dist.8, data = combined_nn, alternative = "greater")
wilcox.test(x = combined_nn$es_dist.9, y = combined_nn$en_dist.9, data = combined_nn, alternative = "greater")
wilcox.test(x = combined_nn$es_dist.10, y = combined_nn$en_dist.10, data = combined_nn, alternative = "greater")
wilcox.test(x = combined_nn$es_dist.11, y = combined_nn$en_dist.11, data = combined_nn, alternative = "greater")
wilcox.test(x = combined_nn$es_dist.12, y = combined_nn$en_dist.12, data = combined_nn, alternative = "greater")
t.test(x = combined_nn$es_dist.1, y = combined_nn$en_dist.1, data = combined_nn)
t.test(x = combined_nn$es_dist.2, y = combined_nn$en_dist.2, data = combined_nn, alternative = "greater")
wilcox.test(x = combined_nn$es_dist.3, y = combined_nn$en_dist.3, data = combined_nn, alternative = "greater")
wilcox.test(x = combined_nn$es_dist.4, y = combined_nn$en_dist.4, data = combined_nn, alternative = "greater")
wilcox.test(x = combined_nn$es_dist.5, y = combined_nn$en_dist.5, data = combined_nn, alternative = "greater")
wilcox.test(x = combined_nn$es_dist.6, y = combined_nn$en_dist.6, data = combined_nn, alternative = "greater")
wilcox.test(x = combined_nn$es_dist.7, y = combined_nn$en_dist.7, data = combined_nn, alternative = "greater")
wilcox.test(x = combined_nn$es_dist.8, y = combined_nn$en_dist.8, data = combined_nn, alternative = "greater")
wilcox.test(x = combined_nn$es_dist.9, y = combined_nn$en_dist.9, data = combined_nn, alternative = "greater")
wilcox.test(x = combined_nn$es_dist.10, y = combined_nn$en_dist.10, data = combined_nn, alternative = "greater")
wilcox.test(x = combined_nn$es_dist.11, y = combined_nn$en_dist.11, data = combined_nn, alternative = "greater")
wilcox.test(x = combined_nn$es_dist.12, y = combined_nn$en_dist.12, data = combined_nn, alternative = "greater")
```

## Fig 3A: Mann-Whitney U tests: is distance of EdU -> Symbiont less than Nuclei -> symbiont

```{r, apo lists, echo = FALSE}
# Comparison between EdU-->symbiont distance and nuclei-->symbiont distance distributions
# EdU to symbiont distance is less than nuclei to symbiont distances
wilcox.test(x = combined_nn$es_dist.1, y = combined_nn$ns_dist.1, data = combined_nn, alternative = "less", conf.int = T)
wilcox.test(x = combined_nn$es_dist.2, y = combined_nn$ns_dist.2, data = combined_nn, alternative = "less", conf.int = T)
wilcox.test(x = combined_nn$es_dist.3, y = combined_nn$ns_dist.3, data = combined_nn, alternative = "less", conf.int = T)
wilcox.test(x = combined_nn$es_dist.4, y = combined_nn$ns_dist.4, data = combined_nn, alternative = "less", conf.int = T)
wilcox.test(x = combined_nn$es_dist.5, y = combined_nn$ns_dist.5, data = combined_nn, alternative = "less", conf.int = T)
wilcox.test(x = combined_nn$es_dist.6, y = combined_nn$ns_dist.6, data = combined_nn, alternative = "less", conf.int = T)
wilcox.test(x = combined_nn$es_dist.7, y = combined_nn$ns_dist.7, data = combined_nn, alternative = "less", conf.int = T)
wilcox.test(x = combined_nn$es_dist.8, y = combined_nn$ns_dist.8, data = combined_nn, alternative = "less", conf.int = T)
wilcox.test(x = combined_nn$es_dist.9, y = combined_nn$ns_dist.9, data = combined_nn, alternative = "less", conf.int = T)
wilcox.test(x = combined_nn$es_dist.10, y = combined_nn$ns_dist.10, data = combined_nn, alternative = "less", conf.int = T)
wilcox.test(x = combined_nn$es_dist.11, y = combined_nn$ns_dist.11, data = combined_nn, alternative = "less", conf.int = T)
wilcox.test(x = combined_nn$es_dist.12, y = combined_nn$ns_dist.12, data = combined_nn, alternative = "less", conf.int = T)

# All highly significant differences between distributions. EdU objects are much further away to symbionts than to host nuclei, as expected. 
# Positive control

# All tests indicate significant location shift in distributions. 
# All tests indicate edu-symbiont distance is less than nuclei-symbiont distance.

```

# Fig 3C: Medians of NN dists, Percentages of EdU+ inside median distance to symbiont

```{r, apo lists, echo = FALSE}

# Find medians of NN distributions for EdU-symbiont and Nuc-symbiont
median(combined_nn$es_dist.1, na.rm = T)
median(combined_nn$ns_dist.1, na.rm = T)
median(combined_nn$es_dist.2, na.rm = T)
median(combined_nn$ns_dist.2, na.rm = T)
median(combined_nn$es_dist.3, na.rm = T)
median(combined_nn$ns_dist.3, na.rm = T)
median(combined_nn$es_dist.4, na.rm = T)
median(combined_nn$ns_dist.4, na.rm = T)
median(combined_nn$es_dist.5, na.rm = T)
median(combined_nn$ns_dist.5, na.rm = T)
median(combined_nn$es_dist.6, na.rm = T)
median(combined_nn$ns_dist.6, na.rm = T)
median(combined_nn$es_dist.7, na.rm = T)
median(combined_nn$ns_dist.7, na.rm = T)
median(combined_nn$es_dist.8, na.rm = T)
median(combined_nn$ns_dist.8, na.rm = T)
median(combined_nn$es_dist.9, na.rm = T)
median(combined_nn$ns_dist.9, na.rm = T)
median(combined_nn$es_dist.10, na.rm = T)
median(combined_nn$ns_dist.10, na.rm = T)
median(combined_nn$es_dist.11, na.rm = T)
median(combined_nn$ns_dist.11, na.rm = T)
median(combined_nn$es_dist.12, na.rm = T)
median(combined_nn$ns_dist.12, na.rm = T)

# Estimate percentage of EdU+ host nuclei inside median distance to symbiont (7 µm) 
count_edu_total <- subset(combined_nn$es_dist.1, combined_nn$es_dist.1 > 0)
count_edu_in <- subset(combined_nn$es_dist.1, combined_nn$es_dist.1 < 7)
count_edu_out <- subset(combined_nn$es_dist.1, combined_nn$es_dist.1 > 7)
perc_edu_in <- length(count_edu_in)/length(count_edu_total)
perc_edu_out <-length(count_edu_out)/length(count_edu_total)
perc_edu_in
perc_edu_out

# Estimate percentage of total host nuclei inside median distance to symbiont (7 µm) 
count_nuc_total <- subset(combined_nn$ns_dist.1, combined_nn$ns_dist.1 > 0)
count_nuc_in <- subset(combined_nn$ns_dist.1, combined_nn$ns_dist.1 < 7)
count_nuc_out <- subset(combined_nn$ns_dist.1, combined_nn$ns_dist.1 > 7)
perc_nuc_in <- length(count_nuc_in)/length(count_nuc_total)
perc_nuc_out <-length(count_nuc_out)/length(count_nuc_total)
perc_nuc_in
perc_nuc_out

# Estimate percentage of total host nuclei inside distance to symbiont (8 µm) 
count_edu_total <- subset(combined_nn$es_dist.1, combined_nn$es_dist.1 > 0)
count_edu_in <- subset(combined_nn$es_dist.1, combined_nn$es_dist.1 < 8)
count_edu_out <- subset(combined_nn$es_dist.1, combined_nn$es_dist.1 > 8)
perc_edu_in <- length(count_edu_in)/length(count_edu_total)
perc_edu_out <-length(count_edu_out)/length(count_edu_total)
perc_edu_in
perc_edu_out

# Estimate percentage of total host nuclei inside distance to symbiont (8 µm) 
count_nuc_total <- subset(combined_nn$ns_dist.1, combined_nn$ns_dist.1 > 0)
count_nuc_in <- subset(combined_nn$ns_dist.1, combined_nn$ns_dist.1 < 8)
count_nuc_out <- subset(combined_nn$ns_dist.1, combined_nn$ns_dist.1 > 8)
perc_nuc_in <- length(count_nuc_in)/length(count_nuc_total)
perc_nuc_out <-length(count_nuc_out)/length(count_nuc_total)
perc_nuc_in
perc_nuc_out


count_edu_total <- subset(combined_nn$es_dist.1, combined_nn$es_dist.1 > 0)
count_edu_in <- subset(combined_nn$es_dist.1, combined_nn$es_dist.1 < 13)
count_edu_out <- subset(combined_nn$es_dist.1, combined_nn$es_dist.1 > 13)
perc_edu_in <- length(count_edu_in)/length(count_edu_total)
perc_edu_out <-length(count_edu_out)/length(count_edu_total)
perc_edu_in
perc_edu_out

count_nuc_total <- subset(combined_nn$ns_dist.1, combined_nn$ns_dist.1 > 0)
count_nuc_in <- subset(combined_nn$ns_dist.1, combined_nn$ns_dist.1 < 13)
count_nuc_out <- subset(combined_nn$ns_dist.1, combined_nn$ns_dist.1 > 13)
perc_nuc_in <- length(count_nuc_in)/length(count_nuc_total)
perc_nuc_out <-length(count_nuc_out)/length(count_nuc_total)
perc_nuc_in
perc_nuc_out

# Estimate percentage of total host nuclei inside distance to symbiont (30 µm) 
count_edu_total <- subset(combined_nn$es_dist.1, combined_nn$es_dist.1 > 0)
count_edu_in <- subset(combined_nn$es_dist.1, combined_nn$es_dist.1 < 30)
count_edu_out <- subset(combined_nn$es_dist.1, combined_nn$es_dist.1 > 30)
perc_edu_in <- length(count_edu_in)/length(count_edu_total)
perc_edu_out <-length(count_edu_out)/length(count_edu_total)
perc_edu_in
perc_edu_out

# Estimate percentage of total host nuclei inside distance to symbiont (30 µm) 
count_nuc_total <- subset(combined_nn$ns_dist.1, combined_nn$ns_dist.1 > 0)
count_nuc_in <- subset(combined_nn$ns_dist.1, combined_nn$ns_dist.1 < 30)
count_nuc_out <- subset(combined_nn$ns_dist.1, combined_nn$ns_dist.1 > 30)
perc_nuc_in <- length(count_nuc_in)/length(count_nuc_total)
perc_nuc_out <-length(count_nuc_out)/length(count_nuc_total)
perc_nuc_in
perc_nuc_out

```

# Fig 3B t-test comparisons

```{r, apo lists, echo = FALSE}

# t-tests describing the means of distributions (previously used Mann-Whitney to find differences between same distributions)
t.test(x = combined_nn$es_dist.1, y = combined_nn$ns_dist.1, data = combined_nn)
t.test(x = combined_nn$es_dist.2, y = combined_nn$ns_dist.2, data = combined_nn)
t.test(x = combined_nn$es_dist.3, y = combined_nn$ns_dist.3, data = combined_nn)
t.test(x = combined_nn$es_dist.4, y = combined_nn$ns_dist.4, data = combined_nn)
t.test(x = combined_nn$es_dist.5, y = combined_nn$ns_dist.5, data = combined_nn)
t.test(x = combined_nn$es_dist.6, y = combined_nn$ns_dist.6, data = combined_nn)
t.test(x = combined_nn$es_dist.7, y = combined_nn$ns_dist.7, data = combined_nn)
t.test(x = combined_nn$es_dist.8, y = combined_nn$ns_dist.8, data = combined_nn)
t.test(x = combined_nn$es_dist.9, y = combined_nn$ns_dist.9, data = combined_nn)
t.test(x = combined_nn$es_dist.10, y = combined_nn$ns_dist.10, data = combined_nn)
t.test(x = combined_nn$es_dist.11, y = combined_nn$ns_dist.11, data = combined_nn)
t.test(x = combined_nn$es_dist.12, y = combined_nn$ns_dist.12, data = combined_nn)

```



## Fig 4E Mann-Whitney U tests: is distance of EdU -> nuclei shorter in apos compared to syms

```{r, apo lists, echo = FALSE}

## Wilcox test for compared NN distances between apos and syms


# Test to show 1st nearest neighbor has a smaller distance than 2nd nearest neighbor.
wilcox.test(x = combined_nn$enapo_dist.1, y = combined_nn$enapo_dist.2, data = combined_nn, alternative = "less")


# Test differences between magnitude of NN distances at k = 1,2,...,12

wilcox.test(x = combined_nn$enapo_dist.1, y = combined_nn$en_dist.1, data = combined_nn, alternative = "less", conf.int = T)
wilcox.test(x = combined_nn$enapo_dist.2, y = combined_nn$en_dist.2, data = combined_nn, alternative = "less", conf.int = T)
wilcox.test(x = combined_nn$enapo_dist.3, y = combined_nn$en_dist.3, data = combined_nn, alternative = "less", conf.int = T)
wilcox.test(x = combined_nn$enapo_dist.4, y = combined_nn$en_dist.4, data = combined_nn, alternative = "less", conf.int = T)
wilcox.test(x = combined_nn$enapo_dist.5, y = combined_nn$en_dist.5, data = combined_nn, alternative = "less", conf.int = T)
wilcox.test(x = combined_nn$enapo_dist.6, y = combined_nn$en_dist.6, data = combined_nn, alternative = "less", conf.int = T)
wilcox.test(x = combined_nn$enapo_dist.7, y = combined_nn$en_dist.7, data = combined_nn, alternative = "less", conf.int = T)
wilcox.test(x = combined_nn$enapo_dist.8, y = combined_nn$en_dist.8, data = combined_nn, alternative = "less", conf.int = T)
wilcox.test(x = combined_nn$enapo_dist.9, y = combined_nn$en_dist.9, data = combined_nn, alternative = "less", conf.int = T)
wilcox.test(x = combined_nn$enapo_dist.10, y = combined_nn$en_dist.10, data = combined_nn, alternative = "less", conf.int = T)
wilcox.test(x = combined_nn$enapo_dist.11, y = combined_nn$en_dist.11, data = combined_nn, alternative = "less", conf.int = T)
wilcox.test(x = combined_nn$enapo_dist.12, y = combined_nn$en_dist.12, data = combined_nn, alternative = "less", conf.int = T)

# Apo distances from edu to nuclei are closer in apo anemones compared to edu-nuclei distances in sym anemones

```
# Fig 4E median comparisons

```{r, apo lists, echo = FALSE}
# Medians for edu-nuclei comparisons for apo and sym
median(combined_nn$enapo_dist.1, na.rm = T)
median(combined_nn$en_dist.1, na.rm = T)
median(combined_nn$enapo_dist.2, na.rm = T)
median(combined_nn$en_dist.2, na.rm = T)
median(combined_nn$enapo_dist.3, na.rm = T)
median(combined_nn$en_dist.3, na.rm = T)
median(combined_nn$enapo_dist.4, na.rm = T)
median(combined_nn$en_dist.4, na.rm = T)
median(combined_nn$enapo_dist.5, na.rm = T)
median(combined_nn$en_dist.5, na.rm = T)
median(combined_nn$enapo_dist.6, na.rm = T)
median(combined_nn$en_dist.6, na.rm = T)
median(combined_nn$enapo_dist.7, na.rm = T)
median(combined_nn$en_dist.7, na.rm = T)
median(combined_nn$enapo_dist.8, na.rm = T)
median(combined_nn$en_dist.8, na.rm = T)
median(combined_nn$enapo_dist.9, na.rm = T)
median(combined_nn$en_dist.9, na.rm = T)
median(combined_nn$enapo_dist.10, na.rm = T)
median(combined_nn$en_dist.10, na.rm = T)
median(combined_nn$enapo_dist.11, na.rm = T)
median(combined_nn$en_dist.11, na.rm = T)
median(combined_nn$enapo_dist.12, na.rm = T)
median(combined_nn$en_dist.12, na.rm = T)
```
# Fig 4E t-test comparisons

```{r, apo lists, echo = FALSE}
# t-tests describing the means of distributions (previously used Mann-Whitney to find differences between same distributions)
t.test(x = combined_nn$enapo_dist.1, y = combined_nn$en_dist.1, data = combined_nn)
t.test(x = combined_nn$enapo_dist.2, y = combined_nn$en_dist.2, data = combined_nn)
t.test(x = combined_nn$enapo_dist.3, y = combined_nn$en_dist.3, data = combined_nn)
t.test(x = combined_nn$enapo_dist.4, y = combined_nn$en_dist.4, data = combined_nn)
t.test(x = combined_nn$enapo_dist.5, y = combined_nn$en_dist.5, data = combined_nn)
t.test(x = combined_nn$enapo_dist.6, y = combined_nn$en_dist.6, data = combined_nn)
t.test(x = combined_nn$enapo_dist.7, y = combined_nn$en_dist.7, data = combined_nn)
t.test(x = combined_nn$enapo_dist.8, y = combined_nn$en_dist.8, data = combined_nn)
t.test(x = combined_nn$enapo_dist.9, y = combined_nn$en_dist.9, data = combined_nn)
t.test(x = combined_nn$enapo_dist.10, y = combined_nn$en_dist.10, data = combined_nn)
t.test(x = combined_nn$enapo_dist.11, y = combined_nn$en_dist.11, data = combined_nn)
t.test(x = combined_nn$enapo_dist.12, y = combined_nn$en_dist.12, data = combined_nn)
```

## plot: Figure 3A

```{r, apo lists, echo = FALSE}

#equal alpha
equal_alpha_plot <- ggplot() +
  geom_density(aes(x=es_dist.1), alpha = 0.8,  fill = "#d95f02", data=edusymfiles2) + 
  geom_density(aes(x=ns_dist.1), alpha = 0.8, fill = "#7570b3", linetype = 6, data=nucsymfiles2) +
  geom_density(aes(x=es_dist.2), alpha = 0.8,  fill = "#d95f02", data=edusymfiles2) +
  geom_density(aes(x=ns_dist.2), alpha = 0.8,  fill = "#7570b3", linetype = 6, data=nucsymfiles2) +
  geom_density(aes(x=es_dist.4), alpha = 0.8,  fill = "#d95f02", data=edusymfiles2) +
  geom_density(aes(x=ns_dist.4), alpha = 0.8,  fill = "#7570b3", linetype = 6, data=nucsymfiles2) +
  geom_density(aes(x=es_dist.6), alpha = 0.8,  fill = "#d95f02", data=edusymfiles2) +
  geom_density(aes(x=ns_dist.6), alpha = 0.8,  fill = "#7570b3", linetype = 6, data=nucsymfiles2) +
  geom_density(aes(x=es_dist.8), alpha = 0.8,  fill = "#d95f02", data=edusymfiles2) +
  geom_density(aes(x=ns_dist.8), alpha = 0.8,  fill = "#7570b3", linetype = 6, data=nucsymfiles2) +
  geom_density(aes(x=es_dist.12), alpha = 0.8,  fill = "#d95f02", data=edusymfiles2) +
  geom_density(aes(x=ns_dist.12), alpha = 0.8,  fill = "#7570b3", linetype = 6, data=nucsymfiles2) + 
  geom_vline(xintercept=7, linetype="dotted", size = .8) +
  xlim(0,100) + theme_minimal() + xlab("nearest neighbor distance (microns)")
#ggsave("equalalpha.pdf", height = 5, width = 6.5)
equal_alpha_plot

#gradient alpha
gradient_plot <- ggplot() +
  geom_density(aes(x=es_dist.1), alpha = 1,  fill = "#d95f02", data=edusymfiles2) + 
  geom_density(aes(x=ns_dist.1), alpha = 1, fill = "#7570b3", linetype = 6, data=nucsymfiles2) +
  geom_density(aes(x=es_dist.2), alpha = 0.9,  fill = "#d95f02", data=edusymfiles2) +
  geom_density(aes(x=ns_dist.2), alpha = 0.9,  fill = "#7570b3", linetype = 6, data=nucsymfiles2) +
  geom_density(aes(x=es_dist.4), alpha = 0.8,  fill = "#d95f02", data=edusymfiles2) +
  geom_density(aes(x=ns_dist.4), alpha = 0.8,  fill = "#7570b3", linetype = 6, data=nucsymfiles2) +
  geom_density(aes(x=es_dist.6), alpha = 0.7,  fill = "#d95f02", data=edusymfiles2) +
  geom_density(aes(x=ns_dist.6), alpha = 0.7,  fill = "#7570b3", linetype = 6, data=nucsymfiles2) +
  geom_density(aes(x=es_dist.8), alpha = 0.6,  fill = "#d95f02", data=edusymfiles2) +
  geom_density(aes(x=ns_dist.8), alpha = 0.6,  fill = "#7570b3", linetype = 6, data=nucsymfiles2) +
  geom_density(aes(x=es_dist.12), alpha = 0.5,  fill = "#d95f02", data=edusymfiles2) +
  geom_density(aes(x=ns_dist.12), alpha = 0.5,  fill = "#7570b3", linetype = 6, data=nucsymfiles2) + 
  geom_vline(xintercept=7, linetype="dotted", size = .8) +
  xlim(0,100) + theme_minimal() + xlab("nearest neighbor distance (microns)")
#ggsave("gradientalpha.pdf", height = 5, width = 6.5)
gradient_plot
```



## figure 3 sym-apo t-test compoarison

```{r, apo lists, echo = FALSE}

apo_sym_mean <- combined_nn %>% summarise_all(funs(mean(., na.rm = TRUE)))
apo_sym_sd <- combined_nn %>% summarise_all(funs(sd(., na.rm = TRUE)))
apo_sym_median <- combined_nn %>% summarise_all(funs(median(., na.rm = TRUE)))
apo_sym_IQR <- combined_nn %>% summarise_all(funs(IQR(., na.rm = TRUE)))
apo_sym_mad <- combined_nn %>% summarise_all(funs(mad(., na.rm = TRUE)))
apo_sym_N <- combined_nn %>% summarise_all(funs(n_distinct(., na.rm = TRUE)))
apo_sym <- bind_rows(apo_sym_mean, apo_sym_sd, apo_sym_median, apo_sym_IQR, apo_sym_mad, apo_sym_N)
apo_sym_sum <- transpose(apo_sym)
colnames(apo_sym_sum) <- c("mean","sd", "median", "IQR", "mad", "n")
rownames(apo_sym_sum) <- colnames(apo_sym)
NNname <- rownames(apo_sym_sum)
apo_sym_sum <- cbind(apo_sym_sum,NNname)
apo_sym_sum_sub <- apo_sym_sum[rownames(apo_sym_sum) %like% "dist.", ]

apo_sym_final <- separate(apo_sym_sum_sub, NNname, c("name","dist","NN"))
apo_sym_final$NN <- as.numeric(apo_sym_final$NN)
apo_sym_final$NN <- as.factor(apo_sym_final$NN)
#apo_sym_final_reordered <- apo_sym_final %>% arrange(NN,name)

```

## plot figure 3B,3C t-test

```{r, apo lists, echo = FALSE}

sym_only <- apo_sym_final[rownames(apo_sym_final) %like% "es|ns" ,]

sym_en_mean_plot <- ggplot(sym_only, aes(x = NN, y=mean, shape = name, color = name, group = name)) +
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), color = "dark gray", width = 0.4, position=position_dodge(width=.5)) +
  geom_point(position=position_dodge(width=.5)) +
  geom_signif(y_position = c(46,52,59,65,70,76,81,86,90,94,97,100), 
              xmin = c(0.75,1.75,2.75,3.75,4.75,5.75,6.75,7.75,8.75,9.75,10.75,11.75),
              xmax = c(1.25,2.25,3.25,4.25,5.25,6.25,7.25,8.25,9.25,10.25,11.25,12.25),
              annotation = c("*"), tip_length = 0,color = "black") +
  theme_minimal() + xlab(expression(paste("Nearest Neighbor (", italic("k"), " 1-12 closest objects)"))) +
  ylab("mean NN distance (microns)") +
  scale_y_continuous(limits = c(-5, 100)) +
  #scale_fill_discrete(name = "State", labels = c("Apo","Sym")) +
  theme(legend.position = c(0.89, 0.12)) +
  labs(color = "State", shape = "State") +
  scale_color_manual(labels = c("EdU+", "Hoechst"), values = c("#d95f02", "#7570b3")) +
  scale_shape_manual(labels = c("EdU+", "Hoechst"), values = c(17, 19))

sym_en_mean_plot 
#ggsave("sym_en_mean_plot.eps", height = 6, width = 9.5)

sym_en_median_plot <- ggplot(sym_only, aes(x = NN, y=median, shape = name, color = name, group = name)) +
  geom_errorbar(aes(ymin=median-IQR, ymax=median+IQR), color = "dark gray", width = 0.4, position=position_dodge(width=.5)) +
  geom_point(position=position_dodge(width=.5)) +
  geom_signif(y_position = c(29,35,43,48,53,57,61,64,68,71,75,79), 
              xmin = c(0.75,1.75,2.75,3.75,4.75,5.75,6.75,7.75,8.75,9.75,10.75,11.75),
              xmax = c(1.25,2.25,3.25,4.25,5.25,6.25,7.25,8.25,9.25,10.25,11.25,12.25),
              annotation = c("*"), tip_length = 0,color = "black") +
  theme_minimal() + xlab(expression(paste("Nearest Neighbor (", italic("k"), " 1-12 closest objects)"))) +
  ylab("median NN distance (microns)") +
  scale_y_continuous(limits = c(0, 100)) +
  #scale_fill_discrete(name = "State", labels = c("Apo","Sym")) +
  theme(legend.position = c(0.89, 0.12)) +
  labs(color = "State", shape = "State") +
  scale_color_manual(labels = c("EdU+", "Hoechst"), values = c("#d95f02", "#7570b3")) +
  scale_shape_manual(labels = c("EdU+", "Hoechst"), values = c(17, 19))

sym_en_median_plot 
#ggsave("sym_en_median_plot.pdf", height = 6, width = 9.5)
#ggsave("sym_en_median_plot.eps", height = 6, width = 9.5)

```

## plot: Figure 4D

```{r, apo lists, echo = FALSE}
# EDU
edu_nuc_nn_plot <- ggplot() +
  geom_density(aes(x=enapo_dist.1), alpha = 0.8,  fill = "pink", data=eduapofiles2) +
  geom_density(aes(x=en_dist.1), alpha = 0.8,  fill = "#1b9e77", data=edunucsfiles2) +
  geom_density(aes(x=enapo_dist.2), alpha = 0.8,  fill = "pink", data=eduapofiles2) +
  geom_density(aes(x=en_dist.2), alpha = 0.8,  fill = "#1b9e77", data=edunucsfiles2) +
  geom_density(aes(x=enapo_dist.4), alpha = 0.8,  fill = "pink", data=eduapofiles2) +
  geom_density(aes(x=en_dist.4), alpha = 0.8,  fill = "#1b9e77", data=edunucsfiles2) +
  geom_density(aes(x=enapo_dist.6), alpha = 0.8,  fill = "pink", data=eduapofiles2) +
  geom_density(aes(x=en_dist.6), alpha = 0.8,  fill = "#1b9e77", data=edunucsfiles2) +
  geom_density(aes(x=enapo_dist.8), alpha = 0.8,  fill = "pink", data=eduapofiles2) +
  geom_density(aes(x=en_dist.8), alpha = 0.8,  fill = "#1b9e77", data=edunucsfiles2) +
  geom_density(aes(x=enapo_dist.12), alpha = 0.8,  fill = "pink", data=eduapofiles2) +
  geom_density(aes(x=en_dist.12), alpha = 0.8,  fill = "#1b9e77", data=edunucsfiles2) +
  xlim(0,35) + theme_minimal() + xlab("Nearest Neighbor distance (microns)")

edu_nuc_nn_plot
#ggsave("edu-nuc-nn.pdf", height = 4, width = 7.5)
```

## plot figure 4E

```{r, apo lists, echo = FALSE}

apo_sym_only <- apo_sym_final[rownames(apo_sym_final) %like% "en", ]
apo_sym_only$name <- as.factor(apo_sym_only$name)
apo_sym_only$name <- factor(apo_sym_only$name,levels=c("enapo","en"))

apo_sym_en_median_plot <- ggplot(apo_sym_only, aes(x = NN, y=median, shape = name, color = name, group = name)) +
  geom_errorbar(aes(ymin=median-mad, ymax=median+mad), color = "dark gray", width = 0.4, position=position_dodge(width=.5)) +
  geom_point(position=position_dodge(width=.5)) +
  geom_signif(y_position = c(9,11,13,14.5,16,17,18,19,20,21,22,23), 
              xmin = c(0.75,1.75,2.75,3.75,4.75,5.75,6.75,7.75,8.75,9.75,10.75,11.75),
              xmax = c(1.25,2.25,3.25,4.25,5.25,6.25,7.25,8.25,9.25,10.25,11.25,12.25),
              annotation = c("*"), tip_length = 0,color = "black") +
  theme_minimal() + xlab(expression(paste("Nearest Neighbor (", italic("k"), " 1-12 closest objects)"))) +
  ylab("median NN distance (microns)") +
  scale_y_continuous(limits = c(0, 25)) +
  #scale_fill_discrete(name = "State", labels = c("Apo","Sym")) +
  theme(legend.position = c(0.85, 0.25)) +
  labs(color = "State", shape = "State") +
  scale_color_manual(labels = c("Apo", "Sym"), values = c("pink","#1b9e77")) +
  scale_shape_manual(labels = c("Apo", "Sym"), values = c(19,17))

apo_sym_en_median_plot
#ggsave("apo_sym_en_median_plot_new.eps", height = 4, width = 8.5)
#ggsave("apo_sym_en_median_plot_new.pdf", height = 4, width = 8.5)
```




