---
title: "Spatial_analyses_symbiont_location"
author: "Trevor Tivey"
date: "1/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#setwd("/local/path/to/data")
set.seed(123)

#libraries
library(data.table)
library(ggplot2)
library(spatstat)
library(plyr)
library(dplyr)
library(tidyr)
library(ggsignif)
library(scales)
scientific_10 <- function(x) {
  parse(text=gsub("e", " %*% 10^", scientific_format()(x)))
  }
```

```{r, echo=FALSE}

### Create Apo Lists

setwd("./apo_analysis/apoout-h-r")
hrfiles <- list.files(pattern = ".txt")
hrfiles
apohrcombined <- adply(hrfiles, 1, read.csv, header = T, sep = "\t")

setwd("../apoout-v-r")
vrfiles <- list.files(pattern = ".txt")
vrfiles
apovrcombined <- adply(vrfiles, 1, read.csv, header = T, sep = "\t")

setwd("../apoout-e-r")
erfiles <- list.files(pattern = ".txt")
erfiles
apoercombined <- adply(erfiles, 1, read.csv, header = T, sep = "\t")

setwd("../..")
#write.csv(apohrcombined, "apo_combined_nuclei.csv")
#write.csv(apoercombined, "apo_combined_edu.csv")
#write.csv(apovrcombined, "apo_combined_volumes.csv")


### Create Sym Lists

setwd("./sym_analysis/nuclei")
hsfiles <- list.files(pattern = ".txt")
hsfiles
hscombined <- adply(hsfiles, 1, read.csv, header = T, sep = "\t")


setwd("../volume")
vsfiles <- list.files(pattern = ".txt")
vsfiles
vscombined <- adply(vsfiles, 1, read.csv, header = T, sep = "\t")


setwd("../edu")
esfiles <- list.files(pattern = ".txt")
esfiles
escombined <- adply(esfiles, 1, read.csv, header = T, sep = "\t")


setwd("../symbionts")
sfiles <- list.files(pattern = ".txt")
sfiles
scombined <- adply(sfiles, 1, read.csv, header = T, sep = "\t")

setwd("../..")
#write.csv(hscombined, "sym_combined_nuclei.csv")
#write.csv(escombined, "sym_combined_edu.csv")
#write.csv(scombined, "sym_combined_symbionts.csv")
#write.csv(vscombined, "sym_combined_volumes.csv")

```

### Summary Statistics

```{r, echo=FALSE}


# apo summary statistics

# hoechst apo
apohrsum <- ddply(apohrcombined, c("X1"), summarise,
                  hN    = length(Volume..pixel.3.),
                  hmean = mean(Volume..pixel.3.),
                  hsd   = sd(Volume..pixel.3.),
                  hse   = hsd / sqrt(hN)
)
apohrsum

# edu apo
apoersum <- ddply(apoercombined, c("X1"), summarise,
                  eN    = length(Volume..micron.3.),
                  emean = mean(Volume..micron.3.),
                  esd   = sd(Volume..micron.3.),
                  ese   = esd / sqrt(eN)
)
apoersum


# sym summary statistics

# hoechst symbiotic
symhssum <- ddply(hscombined, c("X1"), summarise,
                  hN    = length(Volume..pixel.3.),
                  hmean = mean(Volume..pixel.3.),
                  hsd   = sd(Volume..pixel.3.),
                  hse   = hsd / sqrt(hN)
)
symhssum

# edu symbiotic
symessum <- ddply(escombined, c("X1"), summarise,
                  eN    = length(Volume..micron.3.),
                  emean = mean(Volume..micron.3.),
                  esd   = sd(Volume..micron.3.),
                  ese   = esd / sqrt(eN)
)
symessum

# symbiont symbiotic
symsymsum <- ddply(scombined, c("X1"), summarise,
                  sN    = length(Volume..pixel.3.),
                  smean = mean(Volume..pixel.3.),
                  ssd   = sd(Volume..pixel.3.),
                  sse   = ssd / sqrt(sN)
)
symsymsum

```

### Combining Files

```{r, echo=FALSE}

symsymcom <- cbind(vscombined,symsymsum)

symcom <- cbind(vscombined,symessum)
symcom <- cbind(symcom,symhssum)

apocom <- cbind(apovrcombined,apoersum)
apocom <- cbind(apocom,apohrsum)

len_sym <- length(symcom$eN) #95
symcom$state <- rep("sym", len_sym)

len_apo <- length(apocom$eN)
apocom$state <- rep("apo", len_apo)

symapocom <- rbind(symcom,apocom)
```


### Calculate densities

```{r, echo=FALSE}

# Calculate densities
symsymcom$symsymdensity = symsymcom$sN/symsymcom$Volume..micron.3.
symsymcom$symsymdensitymm3 = symsymcom$symsymdensity * 1000000000
mean(symsymcom$symsymdensity)
mean(symsymcom$symsymdensitymm3) # mm3 5.0e5/mm3
sd(symsymcom$symsymdensitymm3) # mm3 2.3e5/mm3
mean(symsymcom$symsymdensitymm3^(2/3)) # mm2 6.4e3/mm2
sd(symsymcom$symsymdensitymm3^(2/3)) # mm2  1.8e3/mm2
range(symsymcom$symsymdensitymm3^(2/3)) # mm2 3.2e3/mm2 to 1.2e4/mm2
182401^(2/3)
1297735^(2/3)

symapocom$edudensity = symapocom$eN/symapocom$Volume..micron.3.
symapocom$edudensitymm3 = symapocom$edudensity * 1000000000
symapocom$edudensity1000 = symapocom$edudensity * 1000000000
symapocom$edudensity2 = symapocom$eN/symapocom$Surface..micron.2.
symapocom$edudensity2mm2 = symapocom$edudensity2 * 1000000

symapocom$hdensity = symapocom$hN/symapocom$Volume..micron.3.
symapocom$hdensitymm3 = symapocom$hdensity * 1000000000

symapocom$percentedu = symapocom$eN/symapocom$hN * 100

symapocom$cellsize = symapocom$Volume..micron.3./symapocom$hN
radiusofhostcell <- (3/(4*pi) * symapocom$cellsize)^(1/3)
symapocom$cellsizediameter = 2*(radiusofhostcell)

```

### T-tests comparing EdU Density, Percent EdU, Hoecsht Density by symbiotic state.
```{r, echo=FALSE}
t.test(edudensity ~ state, data = symapocom) #0.0232
t.test(edudensitymm3 ~ state, data = symapocom) #3.5e5 v 3.9e5
t.test(percentedu ~ state, data = symapocom) #8.747e-5 8.6 v 10.2
t.test(hdensity ~ state, data = symapocom) #0.002189 0.0041 v 0.0038
t.test(hdensitymm3 ~ state, data = symapocom) #4.1e6 v 3.8e6
t.test(cellsizediameter ~ state, data = symapocom) #0.002113 7.80 v 7.95
wilcox.test(edudensity ~ state, data = symapocom) #0.0256
wilcox.test(percentedu ~ state, data = symapocom) #0.0005
wilcox.test(hdensity ~ state, data = symapocom) #0.0015

qqnorm(symapocom$edudensitymm3) 
qqline(symapocom$edudensitymm3) 
qqnorm(symapocom$hdensitymm3) 
qqline(symapocom$hdensitymm3)
qqnorm(symapocom$percentedu)
qqline(symapocom$percentedu)

#apos have higher nuclei density but lower edu density
#apos have lower percent density of edu as well

```

### Summarizing statistics for plotting data

```{r, echo=FALSE}
symapocom_clean <- symapocom[, !duplicated(colnames(symapocom))]

symapocom_ed <- ddply(symapocom_clean, c("state"), summarise,
                      N = length(edudensitymm3),
                      mean = mean(edudensitymm3),
                      sd = sd(edudensitymm3),
                      se = sd / sqrt(N),
                      median = median(edudensitymm3),
                      mad = mad(edudensitymm3)
)
symapocom_ed

symapocom_ep <- ddply(symapocom_clean, c("state"), summarise,
                      N = length(percentedu),
                      mean = mean(percentedu),
                      sd = sd(percentedu),
                      se = sd / sqrt(N),
                      median = median(percentedu),
                      mad = mad(percentedu))
symapocom_ed

symapocom_hd <- ddply(symapocom_clean, c("state"), summarise,
                      N = length(hdensitymm3),
                      mean = mean(hdensitymm3),
                      sd = sd(hdensitymm3),
                      se = sd / sqrt(N),
                      median = median(hdensitymm3),
                      mad = mad(hdensitymm3))
symapocom_hd

```

### Plots comparing (1) EdU density (2) EdU percentage and (3) Nuclei Density

```{r, echo=FALSE}


symapocom$state <- factor(symapocom$state, levels=c("apo","sym"),
                          labels=c("Apo","Sym"))

ggplot(data = symapocom, aes(x = state, y = edudensitymm3, fill = state)) +
  geom_boxplot(aes(fill = state)) +
  scale_y_continuous(label= function(x) {
    ifelse(x==0, "0", parse(text=gsub("[+]", "", gsub("e", " %*% 10^", scientific_format()(x)))))
    },limits = c(0, 1000000),breaks = seq(0,1000000,100000)) + 
  theme_minimal() +
  theme(axis.text.x  = element_text(angle=0, size=10 ), legend.position="none") +
  scale_fill_manual(labels=c("Apo","Sym"),values=c("pink", "#1b9e77")) +
  labs(x="State",y = expression(Proliferative ~ nuclei ~ density ~   ~ (nuclei ~ mm^{-3}))) 
#ggsave("edudensity_fig3.pdf", height = 5, width = 2)

ggplot(data = symapocom, aes(x = state, y = percentedu, fill = state)) + 
  geom_boxplot(aes(group = state, fill = state)) +
  scale_y_continuous(limits = c(0, 25),breaks = seq(0,25,5)) + 
  theme_minimal() +
  theme(axis.text.x  = element_text(angle=0, size=10 ), legend.position="none") +
  scale_fill_manual(labels=c("Apo","Sym"),values=c("pink", "#1b9e77")) +
  labs(x="State",y = expression("% " ~ Proliferative ~ nuclei)) 
#ggsave("edupercent_fig3b.pdf", height = 5, width = 2)

ggplot(data = symapocom, aes(x = state, y = hdensitymm3)) + 
  geom_boxplot(aes(group = state, fill = state)) +
  scale_y_continuous(label= function(x) {
    ifelse(x==0, "0", parse(text=gsub("[+]", "", gsub("e", " %*% 10^", scientific_format()(x)))))
    }, limits = c(0, 8000000),breaks = seq(0,5e7,1e6)) + 
  theme_minimal() +
  theme(axis.text.x  = element_text(angle=0, size=10 ), legend.position="none") +
  scale_fill_manual(labels=c("Apo","Sym"),values=c("pink", "#1b9e77")) +
  labs(x="State",y = expression(Total ~ nuclei ~ density ~   ~ (nuclei ~ mm^{-3}))) 

#ggsave("hdensity_fig3c.pdf", height = 5, width = 2)


```
### Find where nuclei are located on z-axis to determine whether gastrodermal or epidermal

```{r, echo=FALSE}

#escombined
#hscombined
#scombined

#apoercombined
#apohrcombined
#stat summary on Z for each tentacle imaged.

#stat summary on 

zhssum <- ddply(hscombined, c("X1"), summarise,
                hN    = length(Z),
                hmean = mean(Z),
                hsd   = sd(Z),
                hse   = hsd / sqrt(hN),
                hmedian = median(Z)
)


zssum <- ddply(scombined, c("X1"), summarise,
               sN    = length(Z),
               smean = mean(Z),
               ssd   = sd(Z),
               sse   = ssd / sqrt(sN),
               smedian = median(Z)
               
)


zessum <- ddply(escombined, c("X1"), summarise,
                eN    = length(Z),
                emean = mean(Z),
                esd   = sd(Z),
                ese   = esd / sqrt(eN),
                emedian = median(Z)
)


zahsum <- ddply(apohrcombined, c("X1"), summarise,
                ahN    = length(Z),
                ahmean = mean(Z),
                ahsd   = sd(Z),
                ahse   = ahsd / sqrt(ahN),
                ahmedian = median(Z)
)


zaesum <- ddply(apoercombined, c("X1"), summarise,
                aeN    = length(Z),
                aemean = mean(Z),
                aesd   = sd(Z),
                aese   = aesd / sqrt(aeN),
                aemedian = median(Z)
)

```

```{r, echo=FALSE}
# average z-axis location per tentacle

zaesum #apo edu
zahsum #apo hoechst
zessum #sym edu
zssum #sym sym
zhssum #sym hoechst

#test mean z of EdU vs hoechst
t.test(zaesum$aemean, zahsum$ahmean) # 0.1551 no difference, approx 9.5, 9.8
t.test(zessum$emean, zhssum$hmean) # 0.1642 no difference, approx 8.6m 8.3
t.test(zssum$smean, zhssum$hmean) # 0.0545 difference 8.8 8.3
#test difference between edu and hoechst paired per tentacle
t.test(zaesum$aemean, zahsum$ahmean,paired = T) # 2.2e-16 sig difference edu -.3
t.test(zessum$emean, zhssum$hmean, paired = T) # 4.9e-9 sig difference edu + 0.4
t.test(zssum$smean, zhssum$hmean, paired = T) # 3.4e-14 sig difference + 0.5
t.test(zssum$smean, zessum$emean, paired = T) # 0.037 sig difference + 0.1
#for each tentacle, EdU points are closer to mean symbionts than to hoechst 
t.test(zaesum$aemedian, zahsum$ahmedian) # .196 9.3 9.6 enhanced in epidermis in apos

t.test(zaesum$aemedian, zahsum$ahmedian,paired = T) # 7.0e-5 sig difference edu -.3

t.test(zessum$emedian, zhssum$hmedian) # 0.249 7.8 7.5 EdU is found throughout in thicker gastrodermis
t.test(zessum$emedian, zhssum$hmedian, paired = T) # 0.001 sig difference edu + 0.3
t.test(zssum$smedian, zhssum$hmedian, paired = T) # 2.2e-16 sig difference + 1.25
t.test(zssum$smedian, zessum$emedian, paired = T) # 1.7e-10 sig difference + 0.94


#test differences between apo and sym distribution
apodistz = zahsum$ahmean - zaesum$aemean
symdistz = zhssum$hmean - zessum$emean
symdistedusymz = zssum$smean - zessum$emean
apodistmedz = zahsum$ahmedian - zaesum$aemedian
symdistmedz = zhssum$hmedian - zessum$emedian
symdistedusymmedz = zssum$smedian - zessum$emedian
symdistedusymmedz
symdistmedz
symnormalizeddf <- data.frame(symdistz,symdistedusymz, symdistmedz,symdistedusymmedz)


t.test(apodistz,symdistz) # 2.2e-16 different distributions, apo is 
t.test(apodistmedz,symdistmedz) # 8.1e-7 different distributions, apo is 


```
### Create summary dataframe for plotting distribution differences
```{r, echo=FALSE}
prolif_nuc_norm_apo_mean <- zaesum$aemean - zahsum$ahmean
prolif_nuc_norm_apo_med <- zaesum$aemedian - zahsum$ahmedian
prolif_nuc_norm_apo <- data.frame(prolif_nuc_norm_apo_mean, prolif_nuc_norm_apo_med)
prolif_nuc_norm_apo.tidy <- gather(prolif_nuc_norm_apo)
prolif_nuc_norm_sym <- zessum$emean - zhssum$hmean
prolif_nuc_norm_sym_mean <- zessum$emean - zhssum$hmean
prolif_nuc_norm_sym_med <- zessum$emedian - zhssum$hmedian
prolif_nuc_norm_sym <- data.frame(prolif_nuc_norm_sym_mean, prolif_nuc_norm_sym_med)
prolif_nuc_norm_sym.tidy <- gather(prolif_nuc_norm_sym)


combined_normalizeddf <- rbind(prolif_nuc_norm_apo.tidy, prolif_nuc_norm_sym.tidy)
combined_normalizeddf$key <- factor(combined_normalizeddf$key, 
                                    levels=c("prolif_nuc_norm_apo_med","prolif_nuc_norm_sym_med","prolif_nuc_norm_apo_mean","prolif_nuc_norm_sym_mean"))
combined_normalizeddf

combined_normalizeddf_med <- combined_normalizeddf[grep("med", (combined_normalizeddf$key)), ]
combined_normalizeddf_med$key <- factor(combined_normalizeddf_med$key, 
                                    levels=c("prolif_nuc_norm_apo_med","prolif_nuc_norm_sym_med"),
                                    labels=c("Apo","Sym"))

```

```{r, echo=FALSE}

ggplot(data = combined_normalizeddf_med, aes(x = key, y = value)) + 
  geom_boxplot(aes(group = key, fill = key)) +
  scale_y_continuous(limits = c(-3,4),breaks = seq(-3,4,0.5)) + 
  theme_minimal() +
  theme(axis.text.x  = element_text(angle=0, size=10 ), legend.position="none") +
  scale_fill_manual(labels=c("Apo","Sym"), values=c("pink", "#1b9e77")) + 
  xlab("State") + ylab("Normalized Z-axis location") + labs("")
#ggsave("aposymzttest_fig3f.pdf", height = 5, width = 2)


#median
ggplot(symnormalizeddf, aes(x = symdistmedz, y = symdistedusymmedz)) +
  geom_point() +
  geom_density_2d() + theme_minimal() + 
  ylab("Distance from median symbiont center to EdU+ nuclei") +
  xlab("Distance from median host nuclei to EdU+ nuclei") +
  xlim(-4,4) + ylim(-4,4) +
  geom_vline(xintercept=0, size = .8) +
  geom_hline(yintercept=0, size = .8)

#mean
ggplot(symnormalizeddf, aes(x = symdistz, y = symdistedusymz)) +
  geom_point() +
  geom_density_2d() + theme_minimal() + 
  ylab("Distance from median symbiont center to EdU+ nuclei") +
  xlab("Distance from median host nuclei to EdU+ nuclei") +
  xlim(-4,4) + ylim(-4,4) +
  geom_vline(xintercept=0, size = .8) +
  geom_hline(yintercept=0, size = .8)

```
### Finding differences in distributions between apo, sym, hoechst, edu, symbiont combinations
```{r, echo=FALSE}

#escombined
#hscombined
#scombined
#apoercombined
#apohrcombined

t.test(vscombined$Volume..micron.3., apovrcombined$Volume..micron.3.)
t.test(apohrcombined$Z,apoercombined$Z) #apoh 10, apoe 9.6
t.test(escombined$Z,apoercombined$Z) #sym 9.2, apo 9.6
t.test(hscombined$Z,apohrcombined$Z) #sym 8.6, apo 9.9
t.test(escombined$Z,hscombined$Z) #sym 9.2, sym 8.6
t.test(escombined$Z,scombined$Z) #sym 9.2, sym 9.4

wilcox.test(vscombined$Volume..micron.3., apovrcombined$Volume..micron.3.)
wilcox.test(apohrcombined$Z,apoercombined$Z) #apoh 10, apoe 9.6
wilcox.test(escombined$Z,apoercombined$Z) #sym 9.2, apo 9.6
wilcox.test(hscombined$Z,apohrcombined$Z) #sym 8.6, apo 9.9
wilcox.test(escombined$Z,hscombined$Z) #sym 9.2, sym 8.6
wilcox.test(escombined$Z,scombined$Z) #sym 9.2, sym 9.4

wilcox.test(vscombined$Volume..micron.3., apovrcombined$Volume..micron.3.)
wilcox.test(apohrcombined$Z,apoercombined$Z) #apoh 10, apoe 9.6
wilcox.test(escombined$Z,apoercombined$Z) #sym 9.2, apo 9.6
wilcox.test(hscombined$Z,apohrcombined$Z) #sym 8.6, apo 9.9
wilcox.test(escombined$Z,hscombined$Z) #sym 9.2, sym 8.6
wilcox.test(escombined$Z,scombined$Z) #sym 9.2, sym 9.4
wilcox.test(x = combinedfiles2$es_dist.1, y = combinedfiles2$ns_dist.1, data = combinedfiles2, alternative = "less")

```

```{r, echo=FALSE}

ggplot() +
  geom_density(aes(x=ahmedian), alpha = 0.5,  fill = "blue", linetype = 2, data=zahsum) +
  geom_density(aes(x=aemedian), alpha = 0.5,  fill = "yellow", data=zaesum) +
  xlim(0,20) + theme_minimal() + xlab("Z-axis location in tentacle (microns)")
#ggsave("line-ryb-apo-median.pdf", height = 5, width = 6.5)


ggplot() +
  geom_density(aes(x=smedian), alpha = 0.5,  fill = "red", linetype = 3, data=zssum) +
  geom_density(aes(x=hmedian), alpha = 0.5,  fill = "blue", linetype = 2, data=zhssum) +
  geom_density(aes(x=emedian), alpha = 0.5,  fill = "yellow", data=zessum) +
  xlim(0,20) + theme_minimal() + xlab("Z-axis location in tentacle (microns)") 
  
#ggsave("line-ryb-sym-median.pdf", height = 5, width = 6.5)

```

```{r, echo=FALSE}



```

```{r, echo=FALSE}


```

```{r, echo=FALSE}


```

```{r, echo=FALSE}


```



Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
