---
title: "Symbiont_bounding_box_analysis"
author: "Trevor Tivey"
date: "1/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(data.table)
library(plyr)
library(scales)
```

### Step 1:
## Name file lists of symbiont bounding boxes (boxlist) and host total nuclei (pointlist)


```{r, echo=FALSE}
# Set working directory to folder with input files generated from FIJI 3D Object Counter. Files should be lists of symbiont and host objects.
# Symbiont object counter files (suffix = "_2.txt") and host total nuclei object counter files (suffix = " [1-3.txt]")

# Set working directory to folder containing x,y,z coordinates of cell populations
setwd("./sym_analysis/input sym")

# Create subset lists that are organized in corresponding order to each other
# Files containing object coordinates of all symbiont cluster centers of mass (autofluorescence)
boxlist = list.files(pattern = "C1-")
boxlist
# Files containing object coordinates of host EdU-labeled nuclei
edulist = list.files(pattern = " [1-3].txt")
edulist
# Files containing object coordinates of host nuclei (Hoechst labeled)
nuclist = list.files(pattern = "C4-")
nuclist

# Create output filenname list ahead of time (to be written over at the end)
filenm <- sub("txt", "csv", nuclist)
filenm

# pointlist and boxlist samples must correspond to each other. Check the list files to make sure samples are arranged correctly.
nuclist
boxlist

#######
####### Step 2:
####### For loop to determine if (x,y,z) EdU points are within (x,y,z) bounding boxes of symbionts
#######


# For each file read object counter text file as a tab separated csv file.
# points = all total host nuclei in one tentacle
# boxes = all symbiont objects in one tentacle
for(i in 1:length(nuclist)) {
  setwd("./../input sym")
  
  nucpoints <- read.csv(nuclist[i], header = T, sep = "\t")
  boxes <- read.csv(boxlist[i], header = T, sep = "\t")
  
  # define filenm within for loop
  filenm <- sub("txt", "csv", nuclist)
  
  # print to output
  print(filenm[i])
  
  # define X,Y,Z coordinates as numeric

  nucpoints$X = as.numeric(as.character(nucpoints$X))
  nucpoints$Y = as.numeric(as.character(nucpoints$Y))
  nucpoints$Z = as.numeric(as.character(nucpoints$Z))
  
  # define X,Y,Z coordinates of each symbiont bounding box corner as numeric
  boxes$BX = as.numeric(as.character(boxes$BX))
  boxes$BY = as.numeric(as.character(boxes$BY))
  boxes$BZ = as.numeric(as.character(boxes$BZ))
  boxes$B.width = as.numeric(as.character(boxes$B.width))
  boxes$B.height = as.numeric(as.character(boxes$B.height))
  boxes$B.depth = as.numeric(as.character(boxes$B.depth))
  
  # define explicit X,Y,Z coordinates of box by adding width, height and depth to original X,Y,Z coordinate of each box.
  boxes$BXmax <- boxes$BX + boxes$B.width
  boxes$BYmax <- boxes$BY + boxes$B.height
  boxes$BZmax <- boxes$BZ + boxes$B.depth
  
  # turn dataframe into data table
  nucpoints = as.data.table(nucpoints)
  boxes = as.data.table(boxes)
  
  # Create a boolean vector with length of "points": for each point in points, are X,Y,Z coordinates of points found between BX and BXmax, BY and BYmax, BZ and BZmax?
  # It is necessary to do this in one step, if broken up there are many false positives.
  boolw = apply(nucpoints, 1, function(x) {nrow(boxes[(as.numeric(x[11])>=BX) & (as.numeric(x[11])<=BXmax) & (as.numeric(x[12])>=BY) & (as.numeric(x[12])<=BYmax) & (as.numeric(x[13])>=BZ) & (as.numeric(x[13])<=BZmax),])>0})
  nucpoints <- cbind(nucpoints, boolw)
  
  boxes$Bxhalf <- 0.5 * (boxes$B.width / (2^(1/3)))
  boxes$Byhalf <- 0.5 * (boxes$B.height / (2^(1/3)))
  boxes$Bzhalf <- 0.5 * (boxes$B.depth / (2^(1/3)))
  boxes$halfxmin <- boxes$X - boxes$Bxhalf
  boxes$halfxmax <- boxes$X + boxes$Bxhalf
  boxes$halfymin <- boxes$Y - boxes$Byhalf
  boxes$halfymax <- boxes$Y + boxes$Byhalf
  boxes$halfzmin <- boxes$Z - boxes$Bzhalf
  boxes$halfzmax <- boxes$Z + boxes$Bzhalf
  
  boolhalf = apply(nucpoints, 1, function(x) {nrow(boxes[(as.numeric(x[11])>=halfxmin) & (as.numeric(x[11])<=halfxmax) & (as.numeric(x[12])>=halfymin) & (as.numeric(x[12])<=halfymax) & (as.numeric(x[13])>=halfzmin) & (as.numeric(x[13])<=halfzmax),])>0})
  nucpoints <- cbind(nucpoints, boolhalf)
  
  # write new nuclei points file: now contains original nuclei points file with the last column as a boolean vector of whether or not the point is found within a symbiont cluster bounding box.
  setwd("./../nuclei points")
  write.csv(nucpoints, filenm[i]) 
 
}

```

```{r, echo=FALSE}
#edu point list
########################################################################
########################################################################
########################################################################

#######
####### Step 1:
####### Name file lists of symbiont bounding boxes (boxlist) and host total nuclei (pointlist)
#######

# Set working directory to folder with input files generated from FIJI 3D Object Counter. Files should be lists of symbiont and host objects.
# Symbiont object counter files (suffix = "_2.txt") and EdU+ nuclei object counter files (suffix = " [1-3.txt]")

setwd("./sym_analysis/input sym")

filenm <- sub("txt", "csv", edulist)
filenm

boxlist
edulist

# For loop to find if (x,y,z) EdU points are within (x,y,z) bounding boxes of symbionts
for(i in 1:length(edulist)) {
  setwd("./../input sym")
  
  points <- read.csv(edulist[i], header = T, sep = "\t")
  boxes <- read.csv(boxlist[i], header = T, sep = "\t")
  
  filenm <- sub("txt", "csv", edulist)
  print(filenm[i])
  
  points$X = as.numeric(as.character(points$X))
  points$Y = as.numeric(as.character(points$Y))
  points$Z = as.numeric(as.character(points$Z))
  
  boxes$BX = as.numeric(boxes$BX)
  boxes$BY = as.numeric(boxes$BY)
  boxes$BZ = as.numeric(boxes$BZ)
  boxes$B.width = as.numeric(boxes$B.width)
  boxes$B.height = as.numeric(boxes$B.height)
  boxes$B.depth = as.numeric(boxes$B.depth)
  boxes$BXmax <- boxes$BX + boxes$B.width
  boxes$BYmax <- boxes$BY + boxes$B.height
  boxes$BZmax <- boxes$BZ + boxes$B.depth
  
  points = as.data.table(points)
  boxes = as.data.table(boxes)
  boolw = apply(points, 1, function(x) {nrow(boxes[(as.numeric(x[12])>=BX) & (as.numeric(x[12])<=BXmax) & (as.numeric(x[13])>=BY) & (as.numeric(x[13])<=BYmax) & (as.numeric(x[14])>=BZ) & (as.numeric(x[14])<=BZmax),])>0})
  points <- cbind(points, boolw)


  boxes$Bxhalf <- 0.5 * (boxes$B.width / (2^(1/3)))
  boxes$Byhalf <- 0.5 * (boxes$B.height / (2^(1/3)))
  boxes$Bzhalf <- 0.5 * (boxes$B.depth / (2^(1/3)))
  boxes$halfxmin <- boxes$X - boxes$Bxhalf
  boxes$halfxmax <- boxes$X + boxes$Bxhalf
  boxes$halfymin <- boxes$Y - boxes$Byhalf
  boxes$halfymax <- boxes$Y + boxes$Byhalf
  boxes$halfzmin <- boxes$Z - boxes$Bzhalf
  boxes$halfzmax <- boxes$Z + boxes$Bzhalf
  
  boolhalf = apply(points, 1, function(x) {nrow(boxes[(as.numeric(x[12])>=halfxmin) & (as.numeric(x[12])<=halfxmax) & (as.numeric(x[13])>=halfymin) & (as.numeric(x[13])<=halfymax) & (as.numeric(x[14])>=halfzmin) & (as.numeric(x[14])<=halfzmax),])>0})
  points <- cbind(points, boolhalf)
  
  setwd("./../edu points")
  write.csv(points, filenm[i]) 
  
}

```
## INPUT and COMBINED FILES ##
```{r, echo=FALSE}

########################################################################
########################################################################
########################################################################

# Create subset lists that are organized in corresponding order to each other
# Files containing object coordinates of all symbiont cluster centers of mass (autofluorescence)
boxlist = list.files(pattern = "C1-")
boxlist
# Files containing object coordinates of host EdU-labeled nuclei
edulist = list.files(pattern = " [1-3].txt")
edulist
# Files containing object coordinates of host nuclei (Hoechst labeled)
nuclist = list.files(pattern = "C4-")
nuclist


# EdU combined csv file original (no test to see if inside or outside boxes)
setwd("./sym_analysis/input sym")
files <- list.files(pattern = " [1-3].txt")
files
combinededu <- adply(files, 1, read.csv, header = T, sep = "\t")
setwd("./../fig2")
write.csv(combinededu, "combined_edu.csv")

# EdU combined csv file with boolean test
setwd("./../edu points")
files <- list.files(pattern = ".csv")
files
combined <- adply(files, 1, read.csv, header = T)
setwd("./../fig2")
write.csv(combined, "combined.csv")

# Symbiont boxes combined csv file
setwd("./../input sym")
files <- list.files(pattern = "C1-")
files
combinedboxes <- adply(files, 1, read.csv, header = T, sep = "\t")
setwd("./../fig2")
write.csv(combinedboxes, "combined_boxes.csv")

# volumes for each tentacle image combined csv file
setwd("./../volume")
volumelist <- list.files(pattern = ".txt")
volumelist

combinedvolumes <- adply(volumelist, 1, read.csv, header = T, sep = "\t")
setwd("./../fig2")
write.csv(combinedvolumes, "combined_volumes.csv")

# nuclei for each tentacle image combined csv file
setwd("./../input sym")
files <- list.files(pattern = "C4-")
files
combinednuc <- adply(files, 1, read.csv, header = T, sep = "\t")
setwd("./../fig2")
write.csv(combinednuc, "combined_nuclei.csv")


# combined nuclei with booleans combined csv file
setwd("./../nuclei points")
files <- list.files(pattern = "csv")
files
combinednucb <- adply(files, 1, read.csv, header = T)
setwd("./../fig2")
write.csv(combinednucb, "combined_nuclei_bool.csv")

#read them all out again
combinedvolumes <- read.csv("combined_volumes.csv", header = T)
combinedboxes <- read.csv("combined_boxes.csv", header = T)
combined <- read.csv("combined.csv", header = T)
combinednuc <- read.csv("combined_nuclei.csv", header = T)
combinednucb <- read.csv("combined_nuclei_bool.csv", header = T, sep = ",")
```

```{r, echo=FALSE}
# estimated volume per tentacle of overlapping symbiont boxes, from another analysis.

setwd("./sym_analysis/fig2")
combined_redundant <- read.csv("combined_redundant.csv", header = T)
```

```{r, echo=FALSE}

# add column of X*Y*Z volume to symbiont cluster dataframe: combined volumes of symbionts in individual tentacles
combinedboxes$AABBvolume <- combinedboxes$B.width * combinedboxes$B.height * combinedboxes$B.depth
#indtents <- aggregate(Volume..micron.3. ~ X1, combinedboxes, sum)

# create summary data table of total symbiont volumes per tentacle
indtents <- aggregate(AABBvolume ~ X1, combinedboxes, sum)

# create summary data table of total symbiont volume per tentacle, defined directly by aggregated volume of symbiont objects (in pixels^3)
symvol <- aggregate(Volume..pixel.3. ~ X1, combinedboxes, sum)

# summary data table of total symbiont volume per tentacle (in voxels) #same as symvol
symvolvox <- aggregate(Nb.of.obj..voxels ~ X1, combinedboxes, sum)

# add aggregated symbiont AABB volumes to combined volumes dataframe
#combinedvolumesboxes <- cbind(combinedvolumes, indtents$Volume..pixel.3.)
combinedvolumesboxes <- cbind(combinedvolumes, indtents$AABBvolume)
colnames(combinedvolumesboxes)[29] <- "symbiontvolume"
combinedvolumesboxes <- cbind(combinedvolumesboxes, combined_redundant$x)
colnames(combinedvolumesboxes)[30] <- "symbiontcorrection"


combinedvolumesboxes$correctedsymbiontvolume <- combinedvolumesboxes$symbiontvolume - combinedvolumesboxes$symbiontcorrection
#add ratio
combinedvolumesboxes$symaporatio <- combinedvolumesboxes$correctedsymbiontvolume/combinedvolumesboxes$Volume..micron.3.
#add sample names
combinedvolumesboxes <- cbind(volumelist, combinedvolumesboxes)
combinedvolumesboxes <- cbind(combinedvolumesboxes, volumelist)
#subset samples
combinedvolumesboxes <- cbind(combinedvolumesboxes, symvol)
combinedvolumesboxes$symvolrat <- combinedvolumesboxes$Volume..pixel.3./combinedvolumesboxes$Volume..micron.3.


# subset in case symaporatio makes sense. in this case it doesn't need it.
tentaclesubset <- combinedvolumesboxes[combinedvolumesboxes$symaporatio >0,]

```

###  Connecting boolean edu points to tentacles
```{r, echo=FALSE}

########

# aggregate total number of boolean overlaps per sample tentacle
indbools <- aggregate(boolw ~ X1, combined, length)

#select rows containing only true in combined boolean points
comselectedRows <- (combined$boolw == "TRUE")
comselectedRows
combinedreduced <- combined[comselectedRows,]

# number of overlaps (only trues) per tentacle sample
indboolstrue <- aggregate(boolw ~ X1, combinedreduced, length)
indboolstrue

# combined total and true-only lists
indbooltot <- cbind(indbools,indboolstrue$boolw)
indbooltot

#select subset of samples to match tentacle subset from before
selectedRowsind <- (indbooltot$X1 %in% tentaclesubset$X1)
selectedRowsind
indboolred <- indbooltot[selectedRowsind,]
indboolred

#### with half of the sym volume boolhalf
####

# aggregate total number of boolean overlaps per sample tentacle
halfindbools <- aggregate(boolhalf ~ X1, combined, length)

#select rows containing only true in combined boolean points
halfcomselectedRows <- (combined$boolhalf == "TRUE")
halfcomselectedRows
halfcombinedreduced <- combined[halfcomselectedRows,]

# number of overlaps (only trues) per tentacle sample
halfindboolstrue <- aggregate(boolhalf ~ X1, halfcombinedreduced, length)
halfindboolstrue
row24 <- c(24,0)
halfindboolstrueall <- rbind(halfindboolstrue[1:23,], row24, halfindboolstrue[24:94,])

# combined total and true-only lists
halfindbooltot <- cbind(halfindbools,halfindboolstrueall$boolhalf)
halfindbooltot

#select subset of samples to match tentacle subset from before
halfselectedRowsind <- (halfindbooltot$X1 %in% tentaclesubset$X1)
halfselectedRowsind
halfindboolred <- halfindbooltot[halfselectedRowsind,]
halfindboolred

#combined tentacle subset and boolean subsets
tentaclesubsetall <- cbind(tentaclesubset, indboolred[2],indboolred[3], halfindboolred[3])

# rename columns
colnames(tentaclesubsetall)[38] <- "totalpoints"
colnames(tentaclesubsetall)[39] <- "overlappoints"
colnames(tentaclesubsetall)[40] <- "halfoverlappoints"

# add columns for chi-square analysis
tentaclesubsetall$pointratio <- tentaclesubsetall$overlappoints/tentaclesubsetall$totalpoints
tentaclesubsetall$nonoverlappoints <- tentaclesubsetall$totalpoints - tentaclesubsetall$overlappoints
tentaclesubsetall$overlapexpected <- tentaclesubsetall$symvolrat * tentaclesubsetall$totalpoints # redo
tentaclesubsetall$pointspervolume = tentaclesubsetall$totalpoints/tentaclesubsetall$Volume..micron.3.
tentaclesubsetall$observed_expected <- tentaclesubsetall$overlappoints - tentaclesubsetall$overlapexpected # redo
tentaclesubsetall <- cbind(tentaclesubsetall, symvolvox$Nb.of.obj..voxels) #equals the same as symvolume, so instead of comparing to volume, compare to volume voxels
tentaclesubsetall$symvolrat2 <- tentaclesubsetall$Volume..pixel.3./tentaclesubsetall$Nb.of.obj..voxels
```

###  Connect boolean nuclei points to tentacles
```{r, echo=FALSE}

#aggregate total number of boolean overlaps per sample tentacle
indboolsn <- aggregate(boolw ~ X1, combinednucb, length)

#select rows containing only false in combined boolean points
comselectedRowsn <- (combinednucb$boolw == "FALSE")
comselectedRowsn
combinedreducedn <- combinednucb[comselectedRowsn,]

# number of overlaps (only false) per tentacle sample
indboolsnfalse <- aggregate(boolw ~ X1, combinedreducedn, length)
indboolsnfalse

#combined total and true-only lists
indboolntot <- cbind(indboolsn,indboolsnfalse$boolw)
indboolntot

### half boolean
###

#aggregate total number of boolean overlaps per sample tentacle
halfindboolsn <- aggregate(boolhalf ~ X1, combinednucb, length)

#select rows containing only false in combined boolean points
halfcomselectedRowsn <- (combinednucb$boolhalf == "FALSE")
halfcomselectedRowsn
halfcombinedreducedn <- combinednucb[halfcomselectedRowsn,]

# number of overlaps (only false) per tentacle sample
halfindboolsnfalse <- aggregate(boolhalf ~ X1, halfcombinedreducedn, length)
halfindboolsnfalse

#combined total and true-only lists
halfindboolntot <- cbind(halfindboolsn,halfindboolsnfalse$boolhalf)
halfindboolntot

```
###combined tentacle subset and boolean subsets
```{r, echo=FALSE}

tentaclesubsetall <- cbind(tentaclesubsetall, indboolntot[2], indboolntot[3],halfindboolntot[3])

colnames(tentaclesubsetall)[48] <- "totalpointnuc"
colnames(tentaclesubsetall)[49] <- "nonoverlapnuc"
colnames(tentaclesubsetall)[50] <- "halfnonoverlapnuc"
```

### find volumes based on object voxels
```{r, echo=FALSE}
vol_n <- aggregate(Nb.of.obj..voxels ~ X1, combinednucb, sum)

vol_e <- aggregate(Nb.of.obj..voxels ~ X1, combined, sum)

vol_s <- aggregate(Nb.of.obj..voxels ~ X1, combinedboxes, sum)

vol_saabb <- aggregate(AABBvolume ~ X1, combinedboxes, sum)
vols_ens <- data.frame()
vols_ens <- cbind(volumelist,vol_n$Nb.of.obj..voxels,vol_e$Nb.of.obj..voxels,vol_s$Nb.of.obj..voxels, vol_saabb$AABBvolume, combinedvolumes$Nb.of.obj..voxels,combinedvolumes$Volume..micron.3.)
vols_ens <- data.frame(vols_ens)
names(vols_ens) <- c("sample","voln","vole","vols","volsaabb","vol","volmicron")
vols_ens$volsaabb <- as.numeric(as.character(vols_ens$volsaabb))
vols_ens$vols <- as.numeric(as.character(vols_ens$vols))
vols_ens$volsymred <- combined_redundant$x
vols_ens$scorrected <- vols_ens$volsaabb - vols_ens$volsymred
vols_ens$sratio <- vols_ens$volsaabb/vols_ens$vols
vols_ens$spercent <- vols_ens$vols/vols_ens$volsaabb *100
vols_ens$scorrectedratio <- vols_ens$scorrected/vols_ens$vols
vols_ens$scorrectedpercent <- vols_ens$vols/vols_ens$scorrected *100

```

###  Connecting boolean nuclei points to tentacles
```{r, echo=FALSE}
#####

tentaclesubsetall$overlapnuc <- tentaclesubsetall$totalpointnuc - tentaclesubsetall$nonoverlapnuc
tentaclesubsetall$overlapnucexp <- tentaclesubsetall$symvolrat * tentaclesubsetall$totalpointnuc # will change
tentaclesubsetall$nucpervolume = tentaclesubsetall$totalpointnuc/tentaclesubsetall$Volume..micron.3.
tentaclesubsetall$nuc_observed_expected <- tentaclesubsetall$overlapnuc - tentaclesubsetall$overlapnucexp # will change

tentaclesubsetall$nonoverlapexpected <- tentaclesubsetall$totalpoints - tentaclesubsetall$overlapexpected # will change
tentaclesubsetall$chisquare <- (tentaclesubsetall$overlappoints - tentaclesubsetall$overlapexpected)^2/tentaclesubsetall$overlapexpected + 
  (tentaclesubsetall$nonoverlappoints - tentaclesubsetall$nonoverlapexpected)^2/tentaclesubsetall$nonoverlapexpected # will change
tentaclesubsetall$nonoverlapnucexpected <- tentaclesubsetall$totalpointnuc - tentaclesubsetall$overlapnucexp # will change
tentaclesubsetall$chisquarenuc <- (tentaclesubsetall$overlapnuc - tentaclesubsetall$overlapnucexp)^2/tentaclesubsetall$overlapnucexp + 
  (tentaclesubsetall$nonoverlapnuc - tentaclesubsetall$nonoverlapnucexpected)^2/tentaclesubsetall$nonoverlapnucexpected # will change

# wait until after chi square nuc 
tentaclesubsetall$symvolrat3 <- tentaclesubsetall$correctedsymbiontvolume/tentaclesubsetall$Nb.of.obj..voxels
tentaclesubsetall$overlapexpected <- tentaclesubsetall$symvolrat3 * tentaclesubsetall$totalpoints # redo with new ratio
tentaclesubsetall$overlapnucexp <- tentaclesubsetall$symvolrat3 * tentaclesubsetall$totalpointnuc # redo with new ratio
tentaclesubsetall$nuc_observed_expected <- tentaclesubsetall$overlapnuc - tentaclesubsetall$overlapnucexp # redo with new ratio
tentaclesubsetall$observed_expected <- tentaclesubsetall$overlappoints - tentaclesubsetall$overlapexpected # redo with new ratio

tentaclesubsetall$nonoverlapexpected <- tentaclesubsetall$totalpoints - tentaclesubsetall$overlapexpected # redo with new ratio
tentaclesubsetall$chisquare <- (tentaclesubsetall$overlappoints - tentaclesubsetall$overlapexpected)^2/tentaclesubsetall$overlapexpected + 
  (tentaclesubsetall$nonoverlappoints - tentaclesubsetall$nonoverlapexpected)^2/tentaclesubsetall$nonoverlapexpected # redo with new ratio
tentaclesubsetall$nonoverlapnucexpected <- tentaclesubsetall$totalpointnuc - tentaclesubsetall$overlapnucexp # redo with new ratio
tentaclesubsetall$chisquarenuc <- (tentaclesubsetall$overlapnuc - tentaclesubsetall$overlapnucexp)^2/tentaclesubsetall$overlapnucexp + 
  (tentaclesubsetall$nonoverlapnuc - tentaclesubsetall$nonoverlapnucexpected)^2/tentaclesubsetall$nonoverlapnucexpected # redo with new ratio

tentaclesubsetfinal <- subset(tentaclesubsetall[tentaclesubsetall$symvolrat3 < 1,])

tentaclesubsetfinal$csbool <- "False"
tentaclesubsetfinal$csbool[tentaclesubsetfinal$chisquare > 3.84] <- "True"

tentaclesubsetfinal$csbooln <- "False"
tentaclesubsetfinal$csbooln[tentaclesubsetfinal$chisquarenuc > 3.84] <- "True"
```

```{r, echo=FALSE}
##extra half boolean
### observed overlap, observed nonoverlap, expected overlap, expected nonoverlap

# edu halfoverlappoints halfnonoverlapnuc
tentaclesubsetfinal$edu_nonoverlap_half <- tentaclesubsetfinal$totalpoints - tentaclesubsetfinal$halfoverlappoints
tentaclesubsetfinal$nuc_overlap_half <- tentaclesubsetfinal$totalpointnuc - tentaclesubsetfinal$halfnonoverlapnuc

#expected overlaps
tentaclesubsetfinal$halfsymvolrat3 <- tentaclesubsetfinal$symvolrat3 / 2
tentaclesubsetfinal$edu_overlap_exp <- tentaclesubsetfinal$totalpoints * tentaclesubsetfinal$halfsymvolrat3
tentaclesubsetfinal$edu_nonoverlap_exp <- tentaclesubsetfinal$totalpoints - tentaclesubsetfinal$edu_overlap_exp
tentaclesubsetfinal$nuc_overlap_exp <- tentaclesubsetfinal$totalpointnuc * tentaclesubsetfinal$halfsymvolrat3
tentaclesubsetfinal$nuc_nonoverlap_exp <- tentaclesubsetfinal$totalpointnuc - tentaclesubsetfinal$nuc_overlap_exp

tentaclesubsetfinal$halfchisquare <- (tentaclesubsetfinal$halfoverlappoints - tentaclesubsetfinal$edu_overlap_exp)^2/tentaclesubsetfinal$edu_overlap_exp + 
  (tentaclesubsetfinal$edu_nonoverlap_half - tentaclesubsetfinal$edu_nonoverlap_exp)^2/tentaclesubsetfinal$edu_nonoverlap_exp # will change

tentaclesubsetfinal$halfchisquarenuc <- (tentaclesubsetfinal$nuc_overlap_half - tentaclesubsetfinal$nuc_overlap_exp)^2/tentaclesubsetfinal$nuc_overlap_exp + 
  (tentaclesubsetfinal$halfnonoverlapnuc - tentaclesubsetfinal$nuc_nonoverlap_exp)^2/tentaclesubsetfinal$nuc_nonoverlap_exp # will change


tentaclesubsetall$halfchisquarenuc <- (tentaclesubsetall$overlapnuc - tentaclesubsetall$overlapnucexp)^2/tentaclesubsetall$overlapnucexp + 
  (tentaclesubsetall$nonoverlapnuc - tentaclesubsetall$nonoverlapnucexpected)^2/tentaclesubsetall$nonoverlapnucexpected # will change

```
### graphs
```{r, echo=FALSE}
########  graphs
########

library(ggplot2)

# symbiont density on x-axis compared to observed - expected edu+ nuclei
ggplot(data=tentaclesubsetfinal, aes(x=symvolrat3,observed_expected)) + geom_point() + geom_smooth(method = 'glm') +
  theme_classic() + labs(y = "# of EdU+ host nuclei within symbiont clusters \n (observed - expected)", x = "ratio of symbiont density \n(total volume of symbionts / total volume of tentacle)")

# host nuclei on x-axis compared to observed - expected edu+ nuclei
ggplot(data=tentaclesubsetfinal, aes(x=symvolrat3,nuc_observed_expected)) + geom_point() + geom_smooth(method = 'glm') +
  theme_classic() + labs(y = "Host nuclei in symbiont clusters \n (observed - expected)", x = "ratio of symbiont density ratio of tentacle")


ggplot(data=tentaclesubsetfinal, aes(x=symvolrat3,observed_expected, color = csbool)) + geom_point() + geom_smooth(method = 'glm') +
  theme_classic() + labs(y = "# of EdU+ host nuclei in symbiont clusters (observed - expected)", x = "symbiont density ratio of tentacle")


ggplot(data=tentaclesubsetfinal, aes(x=symvolrat3,nuc_observed_expected, color = csbooln)) + geom_point() + geom_smooth(method = 'glm') +
  theme_classic() + labs(y = "# of host nuclei in symbiont clusters (observed - expected)", x = "symbiont density ratio of tentacle")


ggplot(data=tentaclesubsetfinal, aes(x=symvolrat3,chisquare)) + geom_point() + geom_smooth(method = 'glm') +
  theme_classic() + labs(y = "Chi Squares of EdU+ host nuclei in symbiont clusters", x = "symbiont density ratio of tentacle")

ggplot(data=tentaclesubsetfinal, aes(x=chisquarenuc,chisquare)) + geom_point() + geom_smooth(method = 'glm') +
  theme_classic() + labs(y = "Chi Squares of EdU+ host nuclei in symbiont clusters", x = "Chi Squares of host nuclei in symbiont clusters")

```
### Write final datasheets
```{r, echo=FALSE}
setwd("./sym_analysis/fig2")
write.csv(tentaclesubsetall, "tentaclesubsetall.csv")
write.csv(tentaclesubsetfinal, "tentaclesubsetfinal.csv")

tentaclesubsetfinal <- read.csv("tentaclesubsetfinal.csv", header = T)
```

###    chi square info and summed chi square tests
```{r, echo=FALSE}
#####
########

#d.o.f = 1 so should be p = 0.05 of 3.84, p = 0.01 of 6.63, p = 0.001 10.8
# only 5 > 0.001 : 1.56, 9.15, 5.96, 0.38, 
# tentacle 48 had 4 lowest numbers, along with tentacle 17 (which had only a couple symbionts). Tentacles from 17 were removed from analysis

# summed chi squared tests of nuclei overlapped with symbiont clusters
on <- sum(tentaclesubsetfinal$overlapnuc)
non <- sum(tentaclesubsetfinal$nonoverlapnuc)
one <- sum(tentaclesubsetfinal$overlapnucexp)
none <- sum(tentaclesubsetfinal$nonoverlapnucexpected)
ntcsqt <- (on - one)^2/one + (non - none)^2/none
ntcsqt
# 1273.212 redo = 10069.68

# summed chi squared tests of EdU+ nuclei overlapped with symbiont clusters
op <- sum(tentaclesubsetfinal$overlappoints)
nop <- sum(tentaclesubsetfinal$nonoverlappoints)
oe <- sum(tentaclesubsetfinal$overlapexpected)
noe <- sum(tentaclesubsetfinal$nonoverlapexpected)
pcsqt <- (op - oe)^2/oe + (nop - noe)^2/noe
pcsqt
# 609.975

# the numbers of EdU+ nuclei and total nuclei are significantly enriched in symbiont clusters

```

```{r, echo=FALSE}
########    more statistics
########

#paired t-test is used since numbers correlate to same samples.
t.test(x = tentaclesubsetfinal$overlappoints, y = tentaclesubsetfinal$overlapexpected, data = tentaclesubsetfinal, paired = T)
# p = 4.725 x 10^-12
t.test(x = tentaclesubsetfinal$overlapnuc, y = tentaclesubsetfinal$overlapnucexp, data = tentaclesubsetfinal, paired = T)
# p < 2.2 e-16 240 difference

#paired t-test is used since numbers correlate to same samples.
t.test(x = tentaclesubsetfinal$halfoverlappoints, y = tentaclesubsetfinal$edu_overlap_exp, data = tentaclesubsetfinal, paired = T)
# p = 0.0002  9 difference
t.test(x = tentaclesubsetfinal$nuc_overlap_half, y = tentaclesubsetfinal$nuc_overlap_exp, data = tentaclesubsetfinal, paired = T)
# p < 2.2 e-16 240 difference



#Symbiont density and observed - expected
symbiontdensityoelm <- lm(symvolrat3 ~ observed_expected, data = tentaclesubsetfinal)
summary(symbiontdensityoelm)
# pr = 3.44e-9, R = 0.3184

symbiontdensityedulm <- lm(symvolrat3 ~ totalpoints, data = tentaclesubsetfinal)
summary(symbiontdensityedulm)
# pr = 0.7988, R = -0.01049

# calculate edu+ density
tentaclesubsetfinal$edudensity <- tentaclesubsetfinal$totalpoints/tentaclesubsetfinal$Nb.of.obj..voxels
tentaclesubsetfinal$nucdensity <- tentaclesubsetfinal$totalpointnuc/tentaclesubsetfinal$Nb.of.obj..voxels

#symbiont volume ratio is associated with edudensity
symbiontdensityedudensitylm <- lm(symvolrat3 ~ edudensity, data = tentaclesubsetfinal)
summary(symbiontdensityedudensitylm)
# pr = 0.03683, R = 0.03736 F = 4.493

#symbiont volume ratio is not associated with nucdensity
symbiontdensitynucdensitylm <- lm(symvolrat3 ~ nucdensity, data = tentaclesubsetfinal)
summary(symbiontdensitynucdensitylm)
# pr = 0.3296, R = -0.0004342
```

# Figure 3E and F
```{r, echo=FALSE}
ggplot(data=tentaclesubsetfinal, aes(x=symvolrat3,edudensity)) + geom_point() + geom_smooth(method = 'glm') +
  theme_classic() + labs(y = "EdU density", x = "ratio of symbiont density \n(total volume of symbionts / total volume of tentacle)")
setwd("./sym_analysis/fig2")

ggplot(data=tentaclesubsetfinal, aes(x=symvolrat3,totalpoints)) + geom_point() + geom_smooth(method = 'glm') +
  theme_classic() + labs(y = "Total EdU points per tentacle", x = "ratio of symbiont density \n(total volume of symbionts / total volume of tentacle)")


ggplot(data=tentaclesubsetfinal, aes(x=symvolrat3,nucdensity)) + geom_point() + geom_smooth(method = 'glm') +
  theme_classic() + labs(y = "Total host nuclei density", x = "ratio of symbiont density \n(total volume of symbionts / total volume of tentacle)")


# Figure 3F
ggplot(data=tentaclesubsetfinal, aes(x=symvolrat3,edudensity)) + geom_point() + geom_smooth(method = 'glm') +
  theme_minimal() + labs(y = "EdU density", x = "Ratio of symbiont density \n(total volume of symbionts / total volume of tentacle)")

# Figure 3F 
ggplot(data=tentaclesubsetfinal, aes(x=symvolrat3,nucdensity)) + geom_point() + geom_smooth(method = 'glm') +
  theme_minimal() + labs(y = "Total host nuclei density", x = "Ratio of symbiont density \n(total volume of symbionts / total volume of tentacle)")

# Figure 3F no axes titles
ggplot(data=tentaclesubsetfinal, aes(x=symvolrat3,edudensity)) + geom_point() + geom_smooth(method = 'glm') +
  theme_minimal() + labs(y = "", x = "") + scale_y_continuous(label = scientific_format(), limits = c(1e-5,1.6e-4), breaks = c(3e-5,6e-5,9e-5,1.2e-4,1.5e-4))

#ggsave("edu_density_x_ratio.pdf", height = 4, width = 4, useDingbats=FALSE)

# Figure 3F no axes titles
ggplot(data=tentaclesubsetfinal, aes(x=symvolrat3,nucdensity)) + geom_point() + geom_smooth(method = 'glm') +
  theme_minimal() + labs(y = "", x = "") + scale_y_continuous(limits = c(4e-4, 9e-4))

#ggsave("hoechst_density_x_ratio.pdf", height = 4, width = 4, useDingbats=FALSE)


# Figure 3E
ggplot(data=tentaclesubsetfinal, aes(x=symvolrat3,overlappoints)) + geom_point() + geom_smooth(method = 'glm') +
  theme_minimal() + labs(y = "", x = "")
#ggsave("overlap_x_ratio.pdf", height = 4, width = 4, useDingbats=FALSE)


# Figure 3E
ggplot(data=tentaclesubsetfinal, aes(x=symvolrat3,observed_expected)) + geom_point() + geom_smooth(method = 'glm') +
  theme_minimal() + labs(y = "", x = "") + scale_y_continuous(limits = c(-40, 80), breaks = c(-40, 0, 40,80))
#ggsave("o-e_x_ratio.pdf", height = 4, width = 4, useDingbats=FALSE)

```
### Symbiont density comparisons using symbiont volume
```{r, echo=FALSE}
# symbiont density compared to number of proliferating cells within symbiont clusters
symbiontdensityolm <- lm(symvolrat3 ~ overlappoints, data = tentaclesubsetfinal)
summary(symbiontdensityolm)
# pr < 2e-16, R = 0.5986

# symbiont density compared to number of proliferating cells within symbiont clusters
symbiontdensityonuclm <- lm(symvolrat3 ~ nuc_observed_expected, data = tentaclesubsetfinal)
summary(symbiontdensityonuclm)
# pr < 2e-16, R = 0.5986

# symbiont density compared to number of proliferating cells outside of symbiont clusters
symbiontdensitynolm <- lm(symvolrat3 ~ nonoverlappoints, data = tentaclesubsetfinal)
summary(symbiontdensitynolm)
# pr < 7.95e-9, R = 0.3134

# compare symbiont density to density of edu or nuclei within symbiont clusters
# number of edu objects in clusters / total space of symbionts (correctedsymbiontvolume)
tentaclesubsetfinal$educlusterdensity <- tentaclesubsetfinal$overlappoints/tentaclesubsetfinal$correctedsymbiontvolume
tentaclesubsetfinal$nucclusterdensity <- tentaclesubsetfinal$overlapnuc/tentaclesubsetfinal$correctedsymbiontvolume

# 
symbiontdensityeduclusterdensity <- lm(symvolrat3 ~ educlusterdensity, data = tentaclesubsetfinal)
summary(symbiontdensityeduclusterdensity)
# negative effect: pr = 0.88, R = -0.011

symbiontdensitynucclusterdensity <- lm(symvolrat3 ~ nucclusterdensity, data = tentaclesubsetfinal)
summary(symbiontdensitynucclusterdensity)
# negative effect: p = 0.38, R = -0.008


# compare symbiont density to density of edu or nuclei outside symbiont clusters
# number of edu objects outside clusters / total space of symbionts (correctedsymbiontvolume)
tentaclesubsetfinal$eduoutclusterdensity <- tentaclesubsetfinal$nonoverlappoints/(tentaclesubsetfinal$Nb.of.obj..voxels - tentaclesubsetfinal$correctedsymbiontvolume)
tentaclesubsetfinal$nucoutclusterdensity <- tentaclesubsetfinal$nonoverlapnuc/(tentaclesubsetfinal$Nb.of.obj..voxels - tentaclesubsetfinal$correctedsymbiontvolume)

# slight negative effects of symbiont density on density of nuclei and proliferating outside symbiont clusters
symbiontdensityeduoutclusterdensity <- lm(symvolrat3 ~ eduoutclusterdensity, data = tentaclesubsetfinal)
summary(symbiontdensityeduoutclusterdensity)
# negative effect: pr = 0.0381, R = 0.04743

symbiontdensitynucclusterdensity <- lm(symvolrat3 ~ nucoutclusterdensity, data = tentaclesubsetfinal)
summary(symbiontdensitynucclusterdensity)
# negative effect: pr = 4.11e-11, R = 0.3886
```
### Fig. 3D
```{r, echo=FALSE}
tentaclesubsetno24anova <- subset(tentaclesubsetfinal, select=c(overlappoints,nonoverlappoints,overlapexpected))
tentaclesubsetno24anova$nonoverlapexpected <- tentaclesubsetno24anova$overlappoints + tentaclesubsetno24anova$nonoverlappoints - tentaclesubsetno24anova$overlapexpected
ts24anovamelt <- melt(tentaclesubsetno24anova)
anovamelt <- lm(value ~ variable, ts24anovamelt)
summary(anovamelt)
melt.aov <- aov(value ~ variable, ts24anovamelt)
summary(melt.aov)
TukeyHSD(melt.aov)
t.test(x = tentaclesubsetfinal$overlappoints, y = tentaclesubsetfinal$overlapexpected, data = tentaclesubsetfinal)
# p = 0.0185 overlap expected and overlap points

qplot(data = ts24anovamelt, x = variable, y = value)

ts_sum <- ddply(ts24anovamelt, "variable", summarise,
                tsmean = mean(value),
                tsmedian = median(value),
                tssd = sd(value),
                tsIQR = IQR(value),
                tsn = length(value)
)
ts_sum$variable <- as.factor(ts_sum$variable)
ts_sum$variable <- factor(ts_sum$variable,levels=c("overlapexpected","overlappoints","nonoverlapexpected","nonoverlappoints"))

qplot(data = ts_sum, x = variable, y = tsmean)
library(ggsignif)
ggplot(ts_sum, aes(x = variable, y=tsmean, shape = variable, color = variable, fill = variable, group = variable)) +
  geom_errorbar(aes(ymin=tsmean-tssd, ymax=tsmean+tssd), color = "dark gray", width = 0.4, position=position_dodge(width=.5)) +
  geom_point(position=position_dodge(width=.5), size = 3) +
  scale_fill_manual(labels = c("overlappoints", "nonoverlappoints","overlapexpected", "nonoverlapexpected"), values = c("#FF555E", "#FF555E", "white", "white")) +
  scale_color_manual(labels = c("overlappoints", "nonoverlappoints","overlapexpected", "nonoverlapexpected"), values = c("black", "black", "black", "black")) +
  scale_shape_manual(labels = c("overlappoints", "nonoverlappoints","overlapexpected", "nonoverlapexpected"), values = c(22, 22, 22, 22)) +
  theme_minimal() +
  geom_signif(y_position = c(170,235), 
              xmin = c(0.8,2.8),
              xmax = c(2.2,4.2),
              annotation = c("*"), tip_length = 0,color = "black") +
  xlab("") + ylab("") + theme(legend.position = "none") 

setwd("./sym_analysis/fig2")
#ggsave("paired-t-tests-exp-obs.pdf", height = 4, width = 3, useDingbats=FALSE)
```

```{r, echo=FALSE}

#paired t-test is used since numbers correlate to same samples.
t.test(x = tentaclesubsetfinal$overlappoints, y = tentaclesubsetfinal$overlapexpected, data = tentaclesubsetfinal, paired = T)
# p = 4.725 x 10^-12


pointlme <- lm(overlapexpected ~ Volume..micron.3. + symaporatio + Volume..micron.3.:symaporatio, data = tentaclesubsetfinal)
pointlmnoe <- lm(overlapexpected ~ Volume..micron.3. + symaporatio, data = tentaclesubsetfinal)
anova(pointlmnoe,pointlme)
summary(pointlmnoe)
summary(pointlme)


pointlm <- lm(totalpoints ~ Volume..micron.3. + symaporatio + Volume..micron.3.:symaporatio, data = tentaclesubsetfinal)
pointlmno <- lm(totalpoints ~ Volume..micron.3. + symaporatio, data = tentaclesubsetfinal)
anova(pointlmno,pointlm)
summary(pointlmno)
summary(pointlm)

qplot(.resid, .fitted, data = pointlm)
qplot(.resid, totalpoints, data = pointlm)


point.aov <- aov(totalpoints ~ Volume..micron.3. + symvolrat3 + Volume..micron.3.:symvolrat3, data = tentaclesubsetfinal)
qplot(.resid, .fitted, data = point.aov)
qplot(.resid, totalpoints, data = point.aov)
require(proto)
#source(url("http://stat512.cwick.co.nz/code/stat_qqline.r"))
qplot(sample = .resid, data = point.aov) #+ stat_qqline() 
summary(point.aov)

obexpoint.aov <- aov(observed_expected ~ Volume..micron.3. + symvolrat3 + Volume..micron.3.:symvolrat3, data = tentaclesubsetfinal)
qplot(.resid, .fitted, data = obexpoint.aov)
qplot(.resid, observed_expected, data = obexpoint.aov)
require(proto)
#source(url("http://stat512.cwick.co.nz/code/stat_qqline.r"))
qplot(sample = .resid, data = obexpoint.aov) #+ stat_qqline() 
summary(obexpoint.aov)
obexpointlm <- lm(observed_expected ~ Volume..micron.3. + symvolrat3 + Volume..micron.3.:symvolrat3, data = tentaclesubsetfinal)
obexpointlmno <- lm(observed_expected ~ Volume..micron.3. + symvolrat3, data = tentaclesubsetfinal)
obexpointlmnv <- lm(observed_expected ~ Volume..micron.3., data = tentaclesubsetfinal)
obexpointlmnr <- lm(observed_expected ~ symvolrat3, data = tentaclesubsetfinal)
summary(obexpointlmnv)
anova(obexpointlmnv, obexpointlmno)
anova(obexpointlmnr, obexpointlmno)
anova(obexpointlmno, obexpointlm)
anova(obexpointlmnr, obexpointlm)

obexpointlmnn <- lm(observed_expected ~ Volume..micron.3. + correctedsymbiontvolume + correctedsymbiontvolume:Volume..micron.3., data = tentaclesubsetfinal)
summary(obexpointlmnn)

summary(obexpointlmno)
summary(obexpointlm)
anova(obexpointlmno,obexpointlm)

# differences between symbiont to total volume ratios 
# compares total volume of symbiont objects in pixels^3 (voxels) to total volume of host tissue in microns. this is not compatible
# symvolrat <- combinedvolumesboxes$Volume..pixel.3./combinedvolumesboxes$Volume..micron.3.

# compares total volume of symbiont objects in pixels^3 (voxels) to total volume of host tissue in voxels.
# this is not closer to correct but does not take into account the AABB method used to calculate symbiont overlap.
# symvolrat2 <- tentaclesubsetall$Volume..pixel.3./tentaclesubsetall$Nb.of.obj..voxels

# compares total volume of symbiont AABB objects in pixels^3 (voxels), corrected by subtracting overlapping boxes
# to total volume of host tissue in voxels. this is closest to the correct ratio.

#combinedvolumesboxes <- cbind(combinedvolumes, indtents$AABBvolume)
#colnames(combinedvolumesboxes)[29] <- "symbiontvolume"
#combinedvolumesboxes <- cbind(combinedvolumesboxes, combined_redundant$x)
#colnames(combinedvolumesboxes)[30] <- "symbiontcorrection"
#combinedvolumesboxes$correctedsymbiontvolume <- combinedvolumesboxes$symbiontvolume - combinedvolumesboxes$symbiontcorrection

#symvolrat3 <- tentaclesubsetall$correctedsymbiontvolume/tentaclesubsetall$Nb.of.obj..voxels

```

```{r, echo=FALSE}

```

```{r, echo=FALSE}

```
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
